### 2024-11-21 paper/paper.qmd
Exploration of CCHS dataset; Manipulation and Model selection
1. cleaned dataset to handle missing values and proper filter to obtain desired observations of individuals in Ontario

2. Modelled cleaned data using linear regression. Performed model reduction using BIC as there are many predictors.

for future:
start writing up about the dataset with detailed exploration of each predictor and response in paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Chris Yong Hong Sen
thanks: "Code and data are available at: [https://github.com/Monoji77/Alcohol_Use_Ontario](https://github.com/Monoji77/Alcohol_Use_Ontario)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r importing libraries}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph
The estimand would be how much time is spent on exercise for each individual.

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview of dataset

The dataset chosen is obtained from Canadian Community Health Survey Annual component 2017-2018. Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
final_model <-
  readRDS(file = ("../models/final_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "Final model" = final_model
  )
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
```

\newpage

# References

### 2024-11-21 scripts/03-clean_data.R

@@ -1,44 +1,137 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: clean data obotained from https://sda-artsci-utoronto-ca.myaccess.library.utoronto.ca/sdaweb/analysis/?dataset=cchs2017
+# Author: Chris Yong Hong Sen
+# Date: 20 November 2024
+# Contact: luke.yong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: knowledge about parsing csv files using tidyverse functinos
+
 
 #### Workspace setup ####
 library(tidyverse)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Download data ####
+raw_data <- read_csv("../data/01-raw_data/raw_data.csv", na='.')
 
-cleaned_data <-
-  raw_data |>
+#### Clean data ####
+
+# perform simple col name change and ontario filtering
+data_intermediate_1 <- raw_data |>
   janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
+  
+  # rename variables for easier interprebility and variable handling
+  rename(
+    # response
+    time_spent_vigorous_exercise = paadvmva,
+    
+    # potential predictors
+    num_alc_drank_12m  = alc_015,
+    bmi_class = hwtdgbcc,
+    has_mood_disorders = ccc_195,
+    age = dhhgage,
+    sex = dhh_sex,
+    illicit_drug_use = drgdvya,
+    highest_educational_attainment = ehg2dvr3,
+    perceived_life_stress = gen_020,
+    perceived_work_stress = gen_025,
+    health_region = geodghr4,
+    personal_income = incdgper,
+    smoked_hundred_cigarettes = smk_020,
+    
+    # variable to filter province to ontario
+    province = geo_prv
   ) |>
[…]

### 2024-11-21 scripts/06-model_data.R

@@ -10,28 +10,106 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
+library(fastDummies)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv") |>
+  filter(time_spent_vigorous_exercise != 0)  |>
+  mutate(#health_region = factor(health_region),
+         time_spent_vigorous_exercise = log(time_spent_vigorous_exercise)) |>
+  select(-province)
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# run this if generated column names are in scientific notation
+# options(scipen=0)
+analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
+  select(-health_region)
+
+
+### Initial linear regression model ####
+#
+# predictors: all + interaction terms
+# interaction term: 
+#    1. age:highest_educational_attainment
+#    2. age:sex
+#    3. sex:personal_income
+
+original_lmmodel <- lm(time_spent_vigorous_exercise ~ . +
+              age:highest_educational_attainment +
+              age:sex +
+              sex:personal_income,
+                          # age+ 
+                          # highest_educational_attainment+
+                          # personal_income+
+                          # smoked_hundred_cigarettes+
+                          # # perceived_work_stress+
+                          # # illicit_drug_use+
+                          # health_region +
+                          # num_alc_drank_12m,
+              data = analysis_data_w_dummies)
+
+# obtain summary plots to check linear regression assumptions
+# 1. normality of sampling distribution
+# 2. homoscedasticity (random variance)
+# 3. independance of residuals 
+plot(original_lmmodel)
+
+# obtain summary plots 
+summary(original_lmmodel)
+
+# obtain histogram of response variable (log)
[…]

### 2024-11-25 paper/paper.qmd
Added data section in final paper, checked for multicollinearity in linear regression model
1. added VIF for model_data.R script to check for multi collinearity which would lead to inflated standard errors in linear regression model -> ensures that model estimates are reliable by checking VIF values for each estimator is less than 10

2. changed name of response to time_spent_on_vigorous_exercise_7d to reflect that data was collected from respondents regarding time spent on exercise 7 days prior to the phone survey

3. added graphs and tables for three potential explanatory variable and respon

@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+format: 
+  pdf:
+    fig-pos: H
 number-sections: true
@@
 library(palmerpenguins)
+library(knitr)
 ```
@@
 Estimand paragraph
-The estimand would be how much time is spent on exercise for each individual.
+The estimand would be the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on strenuous activities among individuals in Ontario who engage in regular exercise.
 
@@
 
-## Overview of dataset
+## Context of dataset
 
-The dataset chosen is obtained from Canadian Community Health Survey Annual component 2017-2018. Our data [@shelter].... Following @tellingstories, we consider...
+The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada.
+The goal of the CCHS survey is to collect comprehensive and reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
+It aims to provide timely, accessible, and flexible data for monitoring health trends, studying small populations and rare characteristics, and addressing emerging health issues, ultimately helping improve the health and well-being of Canadians.
 
-Overview text
+Other datasets explored include sources from World Health Organisation and ParticipACTION.
+World Health Organisation provides aggregated data informing certain indicators of interest such as obesity levels and alcohol consumption.
+This can be found in the data section for Canada[@OtherSource1].
+Also, ParticipACTION provides key statistics regarding proportion of adult population in canada meeting the national guidelines of 150 minutes of moderate to vigorous exercise, as well as how sedentary canadians are [@OtherSource2].
+However, the availability of raw individual level data is absent from these sources.
+Although the aggregated data gives us a sensing about general Canadian health, we are not able to focus our attention to the Ontario population.
+The dataset chosen from Canadian Community Health Survey (CCHS) annual component is ideal as it contains individual level data, where we are able to obtain the province the respondents are from.
 
-## Measurement
+## Original Dataset
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset provides 113,290 individual response from Canadians all throughout the country.
+Surveys were conducted to collect health-related data from respondents.
+The relevant data collected includes...
+
+- Demographics and Identity: 
+Age, gender, marital status, citizenship, Aboriginal identity, visible minority status, immigrant status, language (mother tongue, spoken at home, first official), sexual orientation.
+
+- Health Status and Conditions:
+Chronic diseases, neurological conditions, mental health indicators, perceived physical and mental health, pain, discomfort, activity limitations, and disabilities.
+
+- Health Behaviors and Lifestyle:
+Smoking, drinking, cannabis use, physical activity, helmet use, dietary practices (e.g., fruits, vegetables), breastfeeding, and BMI.
+
+- Health Care Access and Utilization:
+Contact with health professionals, access to health services, immunizations (e.g., flu shots), cancer screenings (e.g., mammograms, pap smears), waiting times, and satisfaction with services.
+
+- Socio-Economic Factors: 
+Educational attainment, work activity, household income, living arrangements, presence of children in the household, food insecurity.
+
+- Life Quality and Perceptions:
+Life satisfaction, life stress, sense of belonging, health-adjusted life expectancy.
+
+- Specialized Health Indicators:
+Two-week disability days, quality and ratings of health care services, functional health status.
+
+- Specific Population Measures:
+Aboriginal group, immigrant duration, visible minority, minority health disparities.
+
+- Geographic and Household Data:
+Geographic location, number of persons or households, household demographics.
+
+This dataset is highly suitable in determining both quantitative factors such as  c qualitative factors
+
+## Measurement {#sec-measurement}
+
+Our goal is to explore the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on exercise among individuals in Ontario who engage in regular exercise.
+The instrument of measurement would be surveys.
+Through the cross-sectional survey conducted by CCHS, we are able to get the number of hours spent on vigorous exercise per individual, our response variable.
+
+According to the questionnaire, respondents were asked in a phone survey to indicate whether they had engaged in any strenuous activity that lasted at least 10 minutes and made them sweat and breathe a little harder in the past 7 days.
+Afterwards, they were asked to indicate the total time spent on those activities.
+To address the concern of recall bias where an individual may not recall the specific timing in engaging with exercise, surveyors sectioned vigorous exercise into commuting, sports, work and volunteering.
+Then within each section, surveyors asked which day (Monday to Sunday) in the past week had they engaged in such exercise that made them sweat and breathe a little harder.
+By doing so, the activities individuals did during the past week would be anchored on the days of the week.
+Following this would be for the respondents to indicate the total number of hours they engaged in each
+section of vigorous exercise for the past 7 days.
+
+In addition, the explanatory variables we are interested in were also obtained in the same survey.
+
+For socio-demographic factors possibly affecting time spent on vigorous exercise, the surveyors asked the respondent for their age (DHH_AGE) from 0 years old to 121 years old), sex (DHH_SEX) being either male or female, personal income obtained from tax forms submitted to the Canada Revenue Agency upon consent from respondent, geographical health region (ADM_Q037) that was provided by respondents when they agreed to the phone survey, and highest educational attainment (EHG2_04) ranging from less than high school diploma to master's degree and above.
+
+For substance abuse, respondents were asked how frequent they consumed alcoholic drinks in the past 12 months (ALC_Q015), ranging from less than once per month to everyday.
+Surveyors also asked respondents for usage of drugs including one-time marijuana or hashish in the past 12 months (DRG_Q005 to DRG_Q075), and whether they have smoked hundred cigarettes in the past 12 months (SMK_Q020).
+The last 2 questions allowed respondents to indicate a simple yes or no flag to express their answer.
+
+For health factors, questions were asked to determine an individual's body mass index ranging from underweight to obese and perceived life/work stress ranging from not at all stressful to extremely stressful.
+Additionally, surveyors asked respondents if they have mood disorders (E.g.
+depression, bipolarism, mania), for which respondents replied with a simple yes or no.
+
+Overall, each entry obtained by respondents assigns quantitative values to each factor and the the response for which we can explore subsequently using linear regression to determine which factors have greater influence in the time spent exercising amongst individuals that are engaged in exercising.
+
+## Response variable: Time spent on vigorous exercise in the past 7 days
+
+From the CCHS dataset, we look at paadvmva variable.
+This variable is continuous with values ranging from 0 to 9902 representing the number of minutes spent on vigorous activity in the past 7 days.
+For easier readability, the variable will be renamed as time_spent_vigorous_exercise_7d.
+As explained in section @sec-measurement, this variable measures the time spent on any activities lasting more than 10 minutes that makes an individual sweat more than they normally would.
+This would include individuals undertaking laborious employment or volunteering efforts, or it could simply include individuals who do cardio and/or strength training.
+
+By using this variable, we are assuming that the routine of an individual from the past 7 days is indicative of their general routine beyond the past 7 days.
+This variable may not account for individuals who only began vigorous sweating routine within the past 7 days and were sedentary before, which may not be helpful if included in subsequent model fitting.
+
+Since we are interested in individuals that are already engaging in strenuous activities, we can filter time_spent_vigorous_exercise_7d to exclude 0.
+After removing such entries, we obtain the following summary statistics as shown in @tbl-summarised_statistics_time_spent_vigorous_exercise_7d.
+The median time spent on vigorous activities the past 7 days would be 300 minutes, which averages out to 43 minutes per day.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-summarised_statistics_time_spent_vigorous_exercise_7d
+#| tbl-cap: Summarised statistics for time spent on vigorous exercise in the past 7 days
+
+original_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+
+data_cleaned_1 <- original_data |>
+  filter(time_spent_vigorous_exercise_7d != 0)
+
+# data_cleaned_1 |>
+#   top_n(10, time_spent_vigorous_exercise_7d) |>
+#   select(time_spent_vigorous_exercise_7d) |>
+#   arrange(desc(time_spent_vigorous_exercise_7d))
+
+kable(data_cleaned_1 |>
+  summarise(`Min.` = min(time_spent_vigorous_exercise_7d),
+            `1st Qu.` = quantile(time_spent_vigorous_exercise_7d, 0.25),
+            Median = median(time_spent_vigorous_exercise_7d),
+            `3rd Qu.` = quantile(time_spent_vigorous_exercise_7d, 0.75),
+            `Max.` = max(time_spent_vigorous_exercise_7d)))
+
+###############
+# function: obtain outlier observations 
+#
+#
+###############
+
+# first_quartile <- quantile(data_cleaned_1$time_spent_vigorous_exercise_7d, 0.25)
+# third_quartile <- quantile(data_cleaned_1$time_spent_vigorous_exercise_7d, 0.75)
+# IQR <- third_quartile - first_quartile
+# outlier_observation <- data_cleaned_1 |>
+#   filter(time_spent_vigorous_exercise_7d < (first_quartile - 1.5*IQR) | 
+#            time_spent_vigorous_exercise_7d > (third_quartile + 1.5*IQR))
+# nrow(outlier_observation)
+# round(nrow(outlier_observation)/nrow(data_cleaned_1) * 100,0)
+```
+
+Interestingly, the maximum time spent on vigorous activities the past 7 days was 9902 minutes, and this averages to 1415 minutes per day, which is equivalent to 23.5 hours in a day.
+This is very unlikely as this suggests that an individual is constantly engaging in physical strenuous activity without sleeping.
+Logically, this suggests the possibility of outlier values.
+By observing @fig-boxplot_time_spent_vigorous_exercise_7d, there are many outlier observations for this variable.
+Roughly 10% of the observed data are considered outliers as they fall outside 1.5 times the interquartile range from the 25% and 75% quartile values.
+We would be performing log transformation which would be discussed in section (INSERT SECTION HERE PLEASE DONT FORGET) which would address these outlier values.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-boxplot_time_spent_vigorous_exercise_7d
+#| fig-cap: Boxplot of time spent on vigorous exercise in the past 7 days
+#boxplot(data_cleaned_1$time_spent_vigorous_exercise_7d, outcol='indianred4')
+ggplot(data_cleaned_1) +
+  geom_boxplot(aes(y=time_spent_vigorous_exercise_7d), 
+               outlier.color = 'indianred4', 
+               outlier.alpha=0.2) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank()) +
+  ylab('') +
+  xlab('Boxplot of time spent on vigorous exercise in the past 7 days')
+```
+
+Subsequently, we would have to remove observations that are very unlikely as they do not represent the general population and could skew the distribution of time spent on vigorous exercise.
+
+By observing the histogram generated in @fig-hist-time_spent_vigorous_exercise_7d, we obtain a severely right skewed response histogram where most observations lie between 0 to 500 minutes.
+For the building of linear models in subsequent sections of the paper, this would not be ideal as linear regression assumes a normal distribution of the response variable, suggesting that a transformation would be required to fit a linear regression model.
+
+```{r response variable exploration}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-hist-time_spent_vigorous_exercise_7d
+#| fig-cap: Histogram of time spent on vigorous exercise the past 7 days
+
+hist(data_cleaned_1$time_spent_vigorous_exercise_7d,
+     main='',
+     xlab = 'time spent on vigorous exercise the past 7 days / min',
+     ylab = 'no. of individuals',
+     freq=TRUE, cex.main=1, las = 1
+    )
+```
+## Interested explanatory variables
+
+There are a total of 13 predictors of interest that is seemingly related to the time spent in vigorous exercise an exercising individual gets.
+
+### Explanatory Variable 1: Age
+This is a variable of interest because age is seemingly correlated to The age groups of respondents were obtained and they exist in bins that spans ages 18 years to 74 years old.
+
+- 3: 18-19 years old
+- 4: 20-24 years old
+- 5: 25-29 years old
+- 6: 30-34 years old
+- 7: 35-39 years old
+- 8: 40-44 years old
+- 9: 45-49 years old
+- 10: 50-54 years old
+- 11: 55-59 years old
+- 12: 60-64 years old
+- 13: 65-69 years old
+- 14: 70-74 years old
+
+From @fig-barplot-age, we can immediately see that most observations fall within age groups 30 to 34 years old (bin 6), and lesser observations fall within age groups 70-74 years old (bin 14).
+```{r age predictor barplot}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot-age
+#| fig-cap: Histogram of time spent on vigorous exercise the past 7 days
+
+# colnames(data_cleaned_1)
+# 
+# unique((data_cleaned_1 |>
+#          arrange(age))$age)
+#   
+highest_freq_age <- data_cleaned_1 |>
+  count(age) |>
+  top_n(1, n) |>
+  pull(age)
+
+lowest_freq_age <- data_cleaned_1 |>
+  count(age) |>
+  top_n(-1, n) |>
+  pull(age)
+
+data_cleaned_1 |>
+  mutate(group_col = case_when(age == highest_freq_age ~ 'indianred4',
+                               age == lowest_freq_age ~ 'skyblue1',
+                               .default = 'darkgrey')) |>
+  ggplot()  +
+  stat_count(aes(x=age, fill = group_col), show.legend=F) +
+  scale_x_continuous(breaks = c(3:14)) +
+  scale_fill_manual(values = c('indianred4' = 'indianred4', 
+                               'skyblue1'= 'skyblue1', 
+                               'darkgrey'='darkgrey')) +
+  theme_minimal() +
+  xlab('Age in years') +
+  ylab('Number of individuals') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
+From @tbl-summarised_statistics_age and @fig-boxplot-age, we are able to see that the the median age bin belongs to age group 40-44 years old.
+We are able to see that there are no outlier values from @fig-boxplot-age, which means that the survey has obtained an acceptable number of respondents across all age bins.
+```{r age predictor summarised statistics}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-summarised_statistics_age
+#| tbl-cap: Summarised statistics of respondent ages
+kable(data_cleaned_1 |>
+  summarise(`Min.` = min(age),
+            `1st Qu.` = quantile(age, 0.25),
+            Median = median(age),
+            `3rd Qu.` = quantile(age, 0.75),
+            `Max.` = max(age)))
+```
+```{r age predictor boxplot}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-boxplot-age
+#| fig-cap: Boxplot of respondent ages
+
+ggplot(data_cleaned_1) +
+  geom_boxplot(aes(y=age), 
+               outlier.color = 'indianred4', 
+               outlier.alpha=0.2) +
+  theme_minimal() +
+  scale_y_continuous(breaks = c(0:15)) +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()) +
+  ylab('respondent ages') +
+  xlab('')
+```
+
+### Explanatory Variable 2: Sex
+
+Sex is a possible influential predictor for total time spent on vigorous activity.
+Historically, men are associated with exercise and involved in more labour intensive roles such as construction which involves heavy lifting.
+According to @womenExerciseWHO, women are less active than men by 5% since 2000.
+However, with the advent of female only establishments such as female gyms, and 32% increase in memberships at fitness and health clubs from 2010 to 2019, it would be interesting to see if the sex of an individual still affects the time spent on vigorous activites among physically active individuals.
+
+According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents).
+Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected ing @fig-barplot-sex.
+```{r sex predictor barplot}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot-sex
+#| fig-cap: Barplot of sex
+data_cleaned_1 |>
+  mutate(sex = factor(sex, levels = c(0,1), labels =c('Female','Male'))) |>
+  ggplot()  +
+  geom_bar(aes(x=sex, fill=sex), show.legend=F) +
+  scale_fill_manual(values = c('Female' = 'indianred4',
+                               'Male' = 'skyblue1')) +
+  theme_minimal() +
+  xlab('Individual\'s sex') +
+  ylab('Number of individuals') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+
+### obtain actual count of sex ###
+# data_cleaned_1 |>
+#   count(sex, name='no_of_respondents') |>
+#   mutate(sex = ifelse(sex == 0, 'female', 'male'))
+```
+
+### Explanatory Variable 3: Highest level of education
+
+In 2023, @EducationAndActiveLifestyle has done a study to show that highest educational attainment may be a leading factor for physical activity.
+Therefore, it would be worth exploring the effects of highest levels of education with the time spent on vigorous activities among the Ontario residents.
+
+According to @tbl-proportion_highest_educational_attainment, respondents who did not graduate secondary school account for only 5% of the population, and respondents who have at least a bachelor's degree account for 72% of the population.
+According to @fig-barplot_highest_educational_attainment, we are immediately alerted to a class imbalance in our dataset.
+
+```{r proportion of respondents, group(highest_educational_attainment)}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-proportion_highest_educational_attainment
+#| tbl-cap: Proportion of respondents grouped by highest educational level
+kable(data_cleaned_1 |>
+  count(highest_educational_attainment, name = 'proportion_of_respondents') |>
+  mutate(proportion_of_respondents = paste0(round(proportion_of_respondents/nrow(data_cleaned_1)*100,0),'%'),
+         highest_educational_attainment = case_when(
+           highest_educational_attainment == 1 ~ 'less than sec sch grad',
+           highest_educational_attainment == 2 ~ 'sec sch grad',
+           highest_educational_attainment == 3 ~ 'university degree and above',
+           .default = '')))
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot_highest_educational_attainment
+#| fig-cap: Barplot of highest_educational_attainment
+data_cleaned_1 |>
+  mutate(highest_educational_attainment = factor(highest_educational_attainment, levels = c(1,2,3), labels =c('less than sec sch grad','sec sch grad', 'university degree and above'))) |>
+  ggplot()  +
+  geom_bar(aes(x=highest_educational_attainment, fill=highest_educational_attainment), show.legend=F) +
+  scale_fill_manual(values = c('less than sec sch grad' = 'orange1',
+                               'sec sch grad' = 'indianred4',
+                               'university degree and above' = 'skyblue1')) +
+  theme_minimal() +
+  xlab('Highest educational attainment') +
+  ylab('Number of individuals') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+
+```
 

### 2024-11-25 scripts/03-clean_data.R

@@ -11,7 +11,7 @@
 library(tidyverse)
 
 #### Download data ####
-raw_data <- read_csv("../data/01-raw_data/raw_data.csv", na='.')
+raw_data <- read_csv("data/01-raw_data/raw_data.csv", na='.')
 
 #### Clean data ####
 
@@ -22,7 +22,7 @@ data_intermediate_1 <- raw_data |>
   # rename variables for easier interprebility and variable handling
   rename(
     # response
-    time_spent_vigorous_exercise = paadvmva,
+    time_spent_vigorous_exercise_7d = paadvmva,
     
     # potential predictors
     num_alc_drank_12m  = alc_015,
@@ -60,7 +60,7 @@ data_intermediate_1 <- raw_data |>
 # NA-inclusive columns:
 #   num_alc_drank_12m; has_mood_disorders; illicit_drug_use; highest_educational_attainment; 
 #   perceived_life_stress; perceived_work_stress; bmi_class; personal_income; 
-#   time_spent_vigorous_exercise; smoked_hundred_cigarettes;
+#   time_spent_vigorous_exercise_7d; smoked_hundred_cigarettes;
 #
 # NOTE: refer to chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://sda-artsci-utoronto-ca.myaccess.library.utoronto.ca/legacy_sda/dli2/cchs/cchs2017/more_doc/CCHS%202017-2018%20PUMF%20Data%20Dictionary.pdf
 #       for documentation of codes for each variable.
@@ -129,9 +129,29 @@ paste0(round(final_no_observations/original_no_observations*100,0), '%')
 #
 ##########################
 sum(duplicated(data_intermediate2))
- 
+
+
+##########################
+#  cleaning description: 
+#
+#  note: 
+##########################
+data_intermediate3 <- data_intermediate2 |>
+  mutate(
+    # convert 'no' entries in mood_disorders to 0
+    has_mood_disorders = ifelse(has_mood_disorders == 2, 0, has_mood_disorders),
+    
+    # convert 'female' coded 2 to 0
+    sex = ifelse(sex == 2, 0, sex),
+    
+    # convert 'never used drugs' from 2 to 0
+    illicit_drug_use = ifelse(illicit_drug_use == 2, 0, illicit_drug_use),
+    
+    # convert 'never smoked 100 cigarettes' from 2 to 0
[…]

### 2024-11-25 scripts/06-model_data.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simple Linear Regression Model
+# Author: Chris Yong Hong Sen 
+# Date: 24 November 2024
+# Contact:luke.yong@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Knowledge about linear regression models, assumptions about sampling distributino
 
 
 #### Workspace setup ####
@@ -13,10 +12,10 @@ library(tidyverse)
 library(fastDummies)
 
 #### Read data ####
-analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv") |>
-  filter(time_spent_vigorous_exercise != 0)  |>
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv") |>
+  filter(time_spent_vigorous_exercise_7d != 0)  |>
   mutate(#health_region = factor(health_region),
-         time_spent_vigorous_exercise = log(time_spent_vigorous_exercise)) |>
+         time_spent_vigorous_exercise_7d = log(time_spent_vigorous_exercise_7d)) |>
   select(-province)
 
 # run this if generated column names are in scientific notation
@@ -24,7 +23,7 @@ analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv") |>
 analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
   select(-health_region)
 
-
+colnames(analysis_data)
 ### Initial linear regression model ####
 #
 # predictors: all + interaction terms
@@ -32,11 +31,14 @@ analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_regi
 #    1. age:highest_educational_attainment
 #    2. age:sex
 #    3. sex:personal_income
+colnames(analysis_data_w_dummies)
 
-original_lmmodel <- lm(time_spent_vigorous_exercise ~ . +
-              age:highest_educational_attainment +
-              age:sex +
-              sex:personal_income,
+analysis_data_w_dummies |>
[…]

### 2024-11-27 paper/paper.qmd
Refining study, cleaning up repo
1. Added shape files for health region maps which would be used in data exploration section of paper.

2. Completed data exploration section by adding remaining variables.

3. Added model section specifying linear regression model.

4. Started discussion section. Have to complete remaining findings.

5.  Added script which plots map data regarding health region and saved static map plot as png and html. Could potentially create interactive plot and save it as HTML and embed in github repo page.

5. Updated LLM usage for paper in readME and categorise text files

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers"
+subtitle: "Women are engaging in lesser physical activities than men.
+Furthermore, individuals who smoked before are more likely to engage in exercise.
+Consumption of stimulant-related drugs could have also increased as well, resulting in increased overall physical activities in Ontario."
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographic and substance abuse.
+Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to exercise.
+Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels.
+These insights highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing valuable information for designing targeted public health interventions to promote physical activity."
 format: 
@@
 
-## Response variable: Time spent on vigorous exercise in the past 7 days
+## Response variable: Time spent on vigorous activity in the past 7 days
 
@@
 
-### Explanatory Variable 1: Age
+### Variable 1: Age
 This is a variable of interest because age is seemingly correlated to The age groups of respondents were obtained and they exist in bins that spans ages 18 years to 74 years old.
@@
 
-### Explanatory Variable 2: Sex
+### Variable 2: Sex {#sec-sex_exploration}
 
@@
 
-### Explanatory Variable 3: Highest level of education
+### Variable 3: Highest level of education
 
@@
 According to @tbl-proportion_highest_educational_attainment, respondents who did not graduate secondary school account for only 5% of the population, and respondents who have at least a bachelor's degree account for 72% of the population.
-According to @fig-barplot_highest_educational_attainment, we are immediately alerted to a class imbalance in our dataset.
+According to @fig-barplot_highest_educational_attainment, there seems to be class imbalance for this variable.
+
+However, as seen from post-secondary institution enrollment data from @statisticaPostSecEnrollment, there has been an increase in total number of Canadians enrolled in post secondary education from 2000 to 2020.
+Accounting for population changes from 2000 [@statCanadaPopulation2000] to 2022 [@statCanadaPopulation2022], there is still a real increase in the number of real enrollments into post secondary education.
+Therefore, it is very possible that this is representative of the Ontario's population where majority of the population has at least post secondary education.
 
@@
            highest_educational_attainment == 2 ~ 'sec sch grad',
-           highest_educational_attainment == 3 ~ 'university degree and above',
+           highest_educational_attainment == 3 ~ 'post sec certification and above',
            .default = '')))
+
+# Obtain real increase in enrolment
+# pop_2000 <- 30750000
+# pop_2022 <- 39566248
+# 
+# rate_of_enrolment_population_2000 <- 1.44*1000000/pop_2000
+# rate_of_enrolment_population_2020 <- 2.19*1000000/pop_2022 
+# 
+# c(rate_of_enrolment_population_2000, rate_of_enrolment_population_2020)
 ```
@@
 
-## Outcome variables
+### Variable 4: Health Region
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+Health region is a categorical variable which tells us the geographical location an individual belongs to defined by Provincial Ministries of Health.
+It could be possible that certain health regions have more funding that are projected towards increasing awareness towards having an active lifestyle.
+Therefore, it is worth exploring the relationship between health region and time spent on vigorous activities.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+There are a total of 34 different health regions in the Ontario province.
+They are encoded using 5 digits.
+Kindly refer to the appendix (INSERT APPENDIX HERE PLEASE DONT FORGET) to refer to all the various health regions.
+@fig-map_health_region shows us the various health regions in Ontario.
+The blue regions shows us health regions for which lesser respondents belong to while red regions show us health regions for which more respondents belong to.
+
+Highest occurring health regions
+
+- 35951: City of Ottawa Health Unit
+- 35953: Peel Regional Health Unit
+- 35960: Simcoe Muskoka District Health Unit
+- 35970: York Regional Health Unit
+- 35995: City of Toronto Health Unit *(highest number of observations)
+
+Lowest occurring health regions
+
+- 35931: Elgin-St Thomas Health Unit
+- 35935: Haliburton, Kawartha, Pine Ridge District Health Unit
+- 35940: Chatham-Kent Health Unit
+- 35942: Lambton Health Unit
+- 35956: Porcupine Health Unit *(lowest number of observations)
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
 #| echo: false
@@
 #| message: false
+#| label: fig-map_health_region
+#| fig-cap: Map of top five and bottom five occuring Ontario health 
+library(tmap)
+
+top_5_regions <- data_cleaned_1 |>
+  count(health_region, name = 'count') |>
+  top_n(5, count) |>
+  arrange(health_region) |>
+  pull(health_region)
+
+bottom_5_regions <- data_cleaned_1 |>
+  count(health_region, name = 'count') |>
+  top_n(-5, count) |>
+  arrange(health_region) |>
+  pull(health_region)
+
+regions_to_keep <- c(top_5_regions, bottom_5_regions)
+
+source("../scripts/08-health_region_map.R")
+
+health_region_map
 ```
 
-Talk way more about it.
+Overall, @fig-barplot_health_region shows us the actual number of observations belonging to the top five and bottom 5 occurring health regions.
+We can see that most respondents come from the City of Ontario Health Unit (Health_Region = 35995) and lesser respondents belong to Porcupine Health Unit (Health_Region = 35956).
 
-## Predictor variables
+```{r Health region map}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot_health_region
+#| fig-cap: Barplot of top and bottom five occuring health regions
 
-Add graphs, tables and text.
+data_cleaned_1 |>
+  # mutate(health_region = case_when(
+  #   !(health_region %in% regions_to_keep) ~ 'other',
+  #   .default = paste0(health_region, ''))) |>
+  filter(health_region %in% regions_to_keep) |>
+  count(health_region, name = 'count') |>
+  mutate(group = ifelse(health_region %in% top_5_regions, 'top 5 health regions', 'bottom 5 health regions')) |>
+  ggplot() +
+  geom_bar(stat='identity', 
+           aes(y = reorder(health_region, count),
+               x = count,
+               fill = group)) +
+  theme_minimal() +
+  scale_fill_manual(values = c('top 5 health regions' = 'indianred4', 
+                               'bottom 5 health regions' = 'skyblue1')) +
+  theme(legend.title = element_blank(),
+        legend.position = 'bottom',
+        legend.direction = 'vertical',
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank()) +
+  guides(fill = guide_legend(reverse = T)) +
+  ylab('Health Region') +
+  xlab('Number of respondents') 
+```
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+### Variable 5: Personal Income
+
+In the dataset, personal income (INCDGPER) is a categorical variable that places an individuals annual income into a range of income values.
+Below summarises the encoded number and the corresponding meaning of each encoded number.
+All income values are measured per year.
+
+- 1: No income or income loss
+- 2: Less than $20,000 
+- 3: \$20,000 to \$39,999
+- 4: \$40,000 to \$59,999
+- 5: \$60,000 to \$79,999
+- 6: \$80,000 or more
+
+According to @fig-barplot_personal_income, most of our respondents (25% of the sample) in Ontario are earning between \$20,000 to \$39,000 as part of their annual personal income.
+It is also worth mentioning that there are very few respondents who are not earning any income in that year.
+They account for less than 0.5% of the sample.
+In the 2018 annual wages report provided by @PersonalIncomeTaxFiling2018, the annual personal income reported from tax filing in Ontario that year was $39,510.
+Therefore, the sampling population's personal income data is representative of Ontario individuals.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot_personal_income
+#| fig-cap: Barplot of personal income
+
+max_inc <- data_cleaned_1 |> 
+  count(personal_income) |> 
+  top_n(1, n) |>
+  pull(personal_income)
+
+min_inc <- data_cleaned_1 |> 
+  count(personal_income) |> 
+  top_n(-1, n) |>
+  pull(personal_income)
+
+data_cleaned_1 |>
+  mutate(
+    group_col = case_when(
+      personal_income == max_inc ~ 'max',
+       personal_income == min_inc ~ 'min',
+      .default = 'others'),
+    personal_income = factor(personal_income, 
+                               levels = c(1:6), 
+                               labels = c('No income or income loss',
+                                          'Less than $20,000',
+                                          '$20,000 to $39,999',
+                                          '$40,000 to $59,999',
+                                          '$60,000 to $79,999',
+                                          '$80,000 or more'))) |>
+  ggplot()  +
+  geom_bar(aes(x=personal_income, fill=group_col), show.legend=F) +
+  scale_fill_manual(values = c('max'= 'indianred4',
+                               'min' = 'skyblue1',
+                               'others' = 'grey')) +
+  theme_minimal() +
+  xlab('Annual Personal Income') +
+  ylab('Number of respondents') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        axis.text.x = element_text(angle=30, 
+                                   hjust=1,
+                                   vjust=1))
+# obtain counts of each income bin
+# data_cleaned_1 |>
+#   count(personal_income) |>
+#   mutate(prop_of_pop = paste0(round(n/nrow(data_cleaned_1)*100, 1), '%'))
+```
+
+### Variable 6: Frequency of alcohol consumption
+
+This variable is a categorical variable with an implied order.
+
+- 1: Less than once a month
+- 2: Once a month
+- 3: two to three times a month
+- 4: once a week
+- 5: two to three times a week
+- 6: four to six times a week
+- 7: every day
+
+From @fig-barplot_frequency_alcohol_consumption, we are able to see that most respondents from this dataset drink about 2-3 times per week (2973 respondents) as seen from the red bar.
+Also, we are seeing fewer respondents drinking about 4-6 times per week (887 respondents) as seen from the blue bar.
+```{r Alcohol frequency bar graph}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot_frequency_alcohol_consumption
+#| fig-cap: Barplot of frequency of alcohol consumption
+max_alc <- data_cleaned_1 |> 
+  count(num_alc_drank_12m) |> 
+  top_n(1, n) |>
+  pull(num_alc_drank_12m)
+
+min_alc <- data_cleaned_1 |> 
+  count(num_alc_drank_12m) |> 
+  top_n(-1, n) |>
+  pull(num_alc_drank_12m)
+
+data_cleaned_1 |>
+  mutate(
+    group_col = case_when(
+      num_alc_drank_12m == max_alc ~ 'max',
+       num_alc_drank_12m == min_alc ~ 'min',
+      .default = 'others'),
+    num_alc_drank_12m = factor(num_alc_drank_12m, 
+                               levels = c(1:7), 
+                               labels = c('<1 alc per month',
+                                          '1 alc per month',
+                                          '2-3 alc per month',
+                                          '1 alc per week',
+                                          '2-3 alc per week',
+                                          '4-6 alc per week',
+                                          'daily alc'))) |>
+  ggplot()  +
+  geom_bar(aes(x=num_alc_drank_12m, fill=group_col), show.legend=F) +
+  scale_fill_manual(values = c('max'= 'indianred4',
+                               'min' = 'skyblue1',
+                               'others' = 'grey')) +
+  theme_minimal() +
+  xlab('Frequency of alcohol consumption') +
+  ylab('Number of respondents') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        axis.text.x = element_text(angle=30, 
+                                   hjust=1,
+                                   vjust=1.3))
+```
+
+### Variable 7: Illicit drug use the past 12 months
+
+This is a binary variable which indicates whether an individual has used drugs in the past 12 months prior to the interview.
+Drugs include marijuana, hashish, cocaine, amphetamines, speed, methamphetamines, crystal meth, ecstasy, hallucinogens, and sniffing glue, gasoline or other solvents.
+If they have attempted any of these drugs, this variable would be flagged as '1' and '0' otherwise.
+
+It would be interesting to see if drug consumption has an effect on exercising.
+@PhysicalActivityAndSubstanceAbuse exerts that individuals who are more active tend to have lower substance abuse, suggesting a correlation between these two.
+Having this variable would be interesting to explore the relationship of an individuals drug usage and their time spent on exercise.
+
+@fig-barplot_illicit_drug_use shows us the number of respondents who have (or haven't) consumed drugs 12 months prior to their interviews.
+We can see that majority of respondents have not consumed drugs in the given timeframe.
+However, from @tbl-illicit_drug_use, we can see that there are still a number of respondents who had consumed drugs in the past 12 months, particularly a fifth of the sampling population.
+According to @SubstanceUseAndAddiction, they estimate that about 21% of canadians, or 6 million people, will meet the criteria for substance use in their lifetime, and this is inclusive of alcohol and that there is a growing opiod crisis in Ontario.
+Therefore the obtained sample is somewhat representative of usage of drugs among the Ontario population.
+```{r Illicit drug use summary table}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-illicit_drug_use
+#| tbl-cap: Summary of illicit drug use the past 12 months
+
+kable(data_cleaned_1 |>
+  count(illicit_drug_use, name = 'num_of_respondents') |>
+  mutate(prop_of_population = paste0(round(num_of_respondents/nrow(data_cleaned_1)*100,0),'%'),
+         illicit_drug_use = ifelse(illicit_drug_use == 0, 
+                                   'never consumed drugs',
+                                   'consumed at least once')))
+```
+
+```{r Illicit drug use barplot}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot_illicit_drug_use
+#| fig-cap: Barplot of illicit drug use the past 12 months
+data_cleaned_1 |>
+  mutate(illicit_drug_use = factor(illicit_drug_use, levels = c(0,1), labels =c('Non-drug user','Drug user'))) |>
+  ggplot()  +
+  geom_bar(aes(x=illicit_drug_use, fill=illicit_drug_use), show.legend=F) +
+  scale_fill_manual(values = c('Non-drug user' = 'indianred4',
+                               'Drug user' = 'skyblue1')) +
+  theme_minimal() +
+  xlab('') +
+  ylab('Number of respondents') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
+### Variable 8: Cigarette consumption
+
+Respondents were asked whether they had smoked more than 100 cigarettes in their lifetimes.
+This variable is therefore binary, with '1' representing respondents that meet the critera and '0' otherwise.
+
+It is well known that excessive smoking has negative health impacts on individuals.
+A cross sectional study conducted by @SmokingAndPhysicalActiveness found that smokers are less likely to be physically active compared to non-smokers.
+Hence, it would be interesting to discover if there is an effect of smoking more than 100 cigarettes on the time spent on exercise.
+
+@fig-barplot-cigarettes shows us that within the sample, the visual difference between the number of respondents who smoked more than 100 cigarettes (blue column) and respondents who smoked lesser than 100 cigarettes are not that big.
+From @tbl-barplot-cigarettes, 57% of respondents have smoked less than 100 cigarettes while the remaining 43% of respondents smoked more than 100 cigarettes.
+As the classes in this variable is almost split half,
+
+```{r cigarette barplot}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-barplot-cigarettes
+#| fig-cap: Barplot of total cigarettes smoked in a respondent's lifetime
+
+data_cleaned_1 |>
+  mutate(smoked_hundred_cigarettes = factor(smoked_hundred_cigarettes, levels = c(0,1), labels =c('Less than 100 cigarettes','More than 100 cigarettes'))) |>
+  ggplot()  +
+  geom_bar(aes(x=smoked_hundred_cigarettes, fill=smoked_hundred_cigarettes), show.legend=F) +
+  scale_fill_manual(values = c('Less than 100 cigarettes' = 'indianred4',
+                               'More than 100 cigarettes' = 'skyblue1')) +
+  theme_minimal() +
+  xlab('') +
+  ylab('Number of respondents') +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
+@fig-pyrmaid-cigarettes shows us cigarette consumption by age bins among out sample population.
+As respondent age increases from the first bin (18 to 19 year old) to the fourth bin (30 to 34 year old), we can see the number of respondents who have smoked more than 100 cigarettes in their lifetime increased from less than 125 respondents to more than 500 respondents (about 3 times more respondents between the two age groups).
+```{r cigarette consumption pyramid figure}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-pyrmaid-cigarettes
+#| tbl-cap: Pyramid of cigarettes consumption in respondent's lifetime by age groups
+
+data_cleaned_1 |>
+      group_by(age) |>
+  count(smoked_hundred_cigarettes, name = 'num_of_respondents') |>
+  mutate(prop_of_population = paste0(round(num_of_respondents/nrow(data_cleaned_1)*100,0),'%'),
+         smoked_hundred_cigarettes = ifelse(smoked_hundred_cigarettes == 0, 
+                                   'Less than 100 cigarettes',
+                                   'More than 100 cigarettes'),
+         dummy = ifelse(smoked_hundred_cigarettes == 'More than 100 cigarettes', num_of_respondents, -num_of_respondents)) |>
+  ggplot() +
+  geom_col(aes(x=age, y=dummy, fill=smoked_hundred_cigarettes)) +
+  scale_x_continuous(breaks = c(3:14), 
+                     labels = c('18 - 19 y/o',
+                                '20 - 24 y/0',
+                                '25 - 29 y/o',
+                                '30 - 34 y/o',
+                                '35 - 39 y/o',
+                                '40 - 44 y/o',
+                                '45 - 49 y/o',
+                                '50 - 54 y/o',
+                                '55 - 59 y/o',
+                                '60 - 64 y/o',
+                                '65 - 69 y/o',
+                                '70 - 74 y/o')) +
+  scale_y_continuous(breaks = seq(-750,750, 250), labels = abs) +
+  scale_fill_manual(values = c('Less than 100 cigarettes' = 'indianred4',
+                           'More than 100 cigarettes' = 'skyblue1')) +
+  xlab('') +
+  ylab('Number of respondents') +
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.title = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank())
+
+### FOR PERSONAL USE, DELETE LATER
+# colnames(analysis_data_w_dummies)
+# analysis_data |>
+#   select(health_region) |>
+#   distinct() |>
+#   arrange(health_region)
+# length(colnames(analysis_data_w_dummies |>
+#   select(health_region_35927:health_region_35995)))
+# data_cleaned_1 |>
+#   count(smoked_hundred_cigarettes)
+```
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our model is to explore which factors affect time spent on physical activities. Therefore a simple exploratory linear regression would be the chosen 
+modelling algorithm to determine it.
+A full model where our response (time_spent_on_physical_activities) would be fitted on all our interested variables outlined in @sec-data-interestedvariables (INSERT THE SECTION  HERE LATER PLEASE DONT FORGET).
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Let $k$ be the individual from CCHS dataset.
+Define $y_k$ as the time spent on physical activities for individual $k$.
+Then $Age_k$, $Sex_k$, $Educ_k$ and $Alc_k$ represents individual i's age, sex, highest educational attainment, and frequency of alcohol consumption the past 12 months respectively.
+Also, $Drug_k$ and $Cigar_k$ would be indicator explanatory variables derived from the binary variables in our dataset representing if individual k had consumed drugs in the past 12 months and if he/she has smoked 100 cigars in their lifetime respectively.
+Finally, $HealthRegion_k=35953$ and $HealthRegion_k=35970$ are also indicator explanatory variables derived from the categorical variable health_region in our dataset, which represents if individual k belongs to Health Region 35953 (Peel Regional Health Unit) and if he/she belongs to Health Region 35970 (York Regional Health Unit).
+Note that the reference with respect to Health Region would be individuals that are not in both health region 35953 and 35970.
+
+Therefore the final model would be...
+
+\begin{align*} 
+y_k  &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Alc_k +\beta_5I*(Drug_k) + \beta_6I*(Cigar_k) + \\ 
+     & \quad\: \beta_7I*(HealthRegion_k=35953) + \beta_8*I(HealthRegion_k=35970) + \epsilon_k
+\end{align*}
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_k|\mu_k, \sigma^2 &\sim \mbox{Normal}(\mu_k, \sigma^2) \tag{1} \\
+\mu_k &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Alc_k + \tag{2} \\
+      & \quad\: \beta_5I*(Drug_k) + \beta_6I*(Cigar_k) + \notag \\ 
+      & \quad\: \beta_7I*(HealthRegion_k=35953) + \notag \\
+      & \quad\: \beta_8*I(HealthRegion_k=35970) \notag \\
+\epsilon_k &\sim \mbox{Normal}(0, \sigma^2) \tag{3}  
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `stats` package which is a base library offered by the R Core Team, specifically the `lm()` function to create our simple linear regression.
+The default parameters of the `lm()` function would be used.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Initially, we fitted a full model containing every variable of interest.
+Since health region is a categorical variable with no implied order, To handle health region, we will be creating dummy variables for each health region, specifically $I*(HealthRegion_k=i)$ for $i \in {health\_region}$, and the reference chosen for health region would be the health_region = 35926 (The District of Algoma HU).
+We define $Income_k$ to be the the personal income of individual k in our dataset.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Then, for $8 \leq c \leq 40$ and $i \in {health\_region}$, our original full model was defined as...
+
+\begin{align*} 
+y_k  &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Income_k +\beta_5Alc_k +\beta_6I*(Drug_k) +  \\ 
+     & \quad\: \beta_7I*(Cigar_k) + \beta_cI*(HealthRegion=i) + \epsilon_k
+\end{align*}
+
+We then conducted model reduction using backward feature elimination using Bayesian Information Criterion (BIC) as our scoring metric to obtain the least number of predictors that minimises the residual errors from the reduced model.
+This model reduction was performed using `stats` package from the @citeR as well.
+The resulting BIC value was obtained from the `lme4` package [@citationBIC].
 
 # Results
+
+The final model with our estimated coefficients are as follows
+
+\begin{align*} 
+y_k  &= 5.84 - 0.03Age_k + 0.23 Sex_k - 0.08Educ_k + 0.04Alc_k + 0.15I*(Drug_k) +  \\ 
+     & \quad\: 0.12I*(Cigar_k) + \beta_cI*(HealthRegion=i) + \epsilon_k
+\end{align*}
 
@@
 #| message: false
-final_model <-
-  readRDS(file = ("../models/final_model.rds"))
+
 ```
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Linear model of time spent on physically taxing activities based on explanatory variables"
 #| warning: false
+
+final_model <-
+  readRDS(file = ("../models/final_model.rds"))
 
@@
   list(
-    "Final model" = final_model
+    "Time spent on physical activity" = final_model
   )
@@
 
+\newpage
+
 # Discussion
 
-## First discussion point {#sec-first-point}
+## Males are more likely to engage in physical activities {#sec-men_physical_activites}
+For the explanatory variable `sex`, the corresponding coefficient from the fitted linear model is 0.232.
+This means that if an individual is a male, this increases the time spent on physical activities by 26%.
+While we were expecting increase in the overall engagement in physical activity among women as explained in @sec-sex_exploration, men are still more likely to engage in physical activities than women.
+In 2018, Global News Canada reported that 31% of women and 25% of men have insufficient physical activity [@citationMenVsWomenExercise].
+Therefore, our linear regression model discovery was not surprising.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+In fact, in a study published in the Journal of American College of Cardiology, they discovered that women who were physically active would experience greater health benefits than men who were also physically active [@citationWomenSurviveBetterExercise].
+Therefore, this means that more can be done to address the gender gap in the engagement of physical activities.
 
-## Second discussion point
+## Ontario residents who have smoked before might be more likely to engage in physical activities.
+
+From elementary school, Ontario students are exposed to health education where schools would educate about the harms of smoking.
+It is well known that smoking has negative effect's on an individual's health.
+Therefore it is unexpected that the linear regression model showed a positive coefficient for the indicator variable for 'having smoked 100 cigarettes' in an individual's lifetime.
+
+The linear model is telling us that if an individual has smoked 100 cigarettes in their lifetime, they are 13% more likely to spend more time being physically active than an individual who has not smoked more than 100 cigarettes in the same timeframe.
+While it may seem contradicting to what we believe, there may be merit to this value.
+
+A team from the school of Public Health Sciences at University of Waterloo had explored patterns and trends for tobbaco use in canada.
+What they found was that 70% of Canadians who had smoked before have quit as of 2020 @citationSmokingPatternsWaterlooUniversity.
+While smoking may have been affected an individual's physical capabilities, it could have also increased awareness regarding the importance of their health.
+Since most canadians who had a history of smoking had quit smoking, it is very possible that they have greater health conscious.
+Coupled with increased government support for programs with the sole purpose of increasing physical activeness in Ontario [@citationGovernmentSupportExercise], it is no longer surprising why individuals who smoked 100 cigarettes in their lifetime are on average 13% more likely to engage in physical activities.
 
@@
 ## Third discussion point
+
+## Possible increase in stimulant use (about drugs)
 

### 2024-11-27 scripts/06-model_data.R

@@ -21,9 +21,11 @@ analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv") |>
 # run this if generated column names are in scientific notation
 # options(scipen=0)
 analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
-  select(-health_region)
+  select(-c(health_region, has_mood_disorders, 
+            perceived_life_stress,
+            perceived_work_stress))
 
-colnames(analysis_data)
+colnames(analysis_data_w_dummies)
 ### Initial linear regression model ####
 #
 # predictors: all + interaction terms
@@ -35,6 +37,7 @@ colnames(analysis_data_w_dummies)
 
 analysis_data_w_dummies |>
   count(personal_income)
+
 original_lmmodel <- lm(time_spent_vigorous_exercise_7d ~ ., #+
               # age:highest_educational_attainment +
               # age:sex +
@@ -58,32 +61,33 @@ plot(original_lmmodel)
 # obtain summary plots 
 summary(original_lmmodel)
 
+
 # obtain histogram of response variable (log)
 hist(analysis_data_w_dummies$time_spent_vigorous_exercise_7d)
 
+
+
+##############################
+# Model scoring metric: F test hypothesis testing
+#
+# rationale: from comparing reduced model by any 1 predictor with 
+#            the full model, we can deduce if the removal of the 
+#            predictor results in difference in regression power
+#            of in the reduced model
+#
+# goal: remove predictors that fail the anova F test, which means 
+#       removing the predictor does not change the power of the 
+#       reduced model. This decreases complexity of linear regression 
+# 
+##############################
 predictors <- colnames(analysis_data_w_dummies |>
   select(-time_spent_vigorous_exercise_7d))
 
 p_value_vec <- c()
 for (pred in predictors) {
+  
   reduced_data <- analysis_data_w_dummies |> select(-all_of(pred)) 
-  
-  # age:highest_educational_attainment +
-  #   age:sex +
-  # sex:personal_income
-  # if (pred == 'age') {
-  #   reduced_model <- lm(time_spent_vigorous_exercise_7d ~ .+sex:personal_income, data = reduced_data)
-  # } else if (pred == 'highest_educational_attainment') {
-  #   reduced_model <- lm(time_spent_vigorous_exercise_7d ~ . + age:sex + sex:personal_income, data = reduced_data)
-  # } else if (pred == 'sex') {
[…]

### 2024-11-27 scripts/08-health_region_map.R

@@ -1,0 +1,65 @@
+
+library(sf)
+library(tmap)
+library(tidyverse)
+health_region_shp <- st_read(
+  "../data/03-map_data/Health Regions Ontario Shapes/HR_035a18a_e.shp",
+  quiet = T)
+
+# tm_shape(canada_shp) +
+#   tm_borders() +
+tmap_mode('plot')
+
+HR_UID_top <- tibble(HU_ID = top_5_regions) |>
+  mutate(health_region = case_when(
+    HU_ID == 35951 ~ 'City of Ottawa Health Unit',
+    HU_ID == 35953 ~ 'Peel Regional Health Unit',
+    HU_ID == 35960 ~ 'Simcoe Muskoka District Health Unit',
+    HU_ID == 35970 ~ 'York Regional Health Unit',
+    HU_ID == 35995 ~ 'City of Toronto Health Unit'
+  ))
+
+HR_UID_btm <- tibble(HU_ID = bottom_5_regions) |>
+  mutate(health_region = case_when(
+    HU_ID == 35931 ~ 'Elgin-St Thomas Health Unit',
+    HU_ID == 35935 ~ 'Haliburton, Kawartha, Pine Ridge District ',
+    HU_ID == 35940 ~ 'Chatham-Kent Health Unit',
+    HU_ID == 35942 ~ 'Lambton Health Unit',
+    HU_ID == 35956 ~ 'Porcupine Health Unit'
+  ))
+
+
+
+HR_UID_top <- paste0(substr(top_5_regions, 1, 2), substr(top_5_regions, 4, 5))
+
+HR_UID_btm <- paste0(substr(bottom_5_regions, 1, 2), substr(bottom_5_regions, 4, 5))
+
+top_health_regions_shp <- health_region_shp |>
+  filter(HR_UID %in% HR_UID_top)
+
+btm_health_regions_shp <- health_region_shp |>
+  filter(HR_UID %in% HR_UID_btm)
+
+
+health_region_map <- tm_basemap('CartoDB.Positron') +
+  tm_shape(health_region_shp) +
+  tm_borders(colour = 'ENGNAME',
+             legend.show = F,
+             popup.vars = c('Name: '='ENGNAME'),
+             id = 'ENGNAME') +
+  tm_shape(top_health_regions_shp) +
+  tm_polygons(fill='indianred4',
+              fill_alpha = 0.5) +
+  tm_shape(btm_health_regions_shp) +
+  tm_polygons(fill='skyblue1',
[…]

### 2024-12-02 paper/paper.qmd
Major update for paper
1. cleaned up paper.qmd file:
 - added diagnostic check in appendix
 - added one more discussion point

2. removed unnecessary files

3. added to model justification regarding BIC values

4. added parquet file for analysis data

5. cleaned references

@@
 subtitle: "Women are engaging in lesser physical activities than men.
-Furthermore, individuals who smoked before are more likely to engage in exercise.
+Furthermore, individuals who smoked before are more likely to engage in physical activity.
 Consumption of stimulant-related drugs could have also increased as well, resulting in increased overall physical activities in Ontario."
@@
 abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographic and substance abuse.
-Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to exercise.
+Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to be physically active.
 Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels.
@@
 
-library(tidyverse)
-library(palmerpenguins)
-library(knitr)
+library(tidyverse) # for manipulation and visualisation
+library(knitr) # displaying tables
+library(arrow) # reading parquet file
+library(tmap) # displaying maps
+library(lme4) # reporting BIC values
 ```
@@
 
-Overview paragraph
-
-Estimand paragraph
-The estimand would be the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on strenuous activities among individuals in Ontario who engage in regular exercise.
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Physical activity is a cornerstone of public health, yet participation rates remain suboptimal across various populations in Ontario.
+This is especially concerning as physical inactivity is a leading contributor to chronic diseases such as heart disease, diabetes, and mental health disorders.
+Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play a critical role in shaping physical activity behaviors.
+Understanding these factors and their complex interactions is crucial for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.
+
+Recent studies suggest that certain demographic groups, such as older adults and individuals with lower educational levels, are more likely to experience barriers to physical activity, while alcohol consumption and smoking are associated with lower physical activity levels [@citationBaumanPhysicalActivity].
+However, limited research has explored how these factors specifically interact in Ontario, a province with diverse populations and health regions.
+As the burden of chronic disease grows, addressing these disparities becomes even more urgent [@citationWarburtonPhysicalActivity] .
+
+This study seeks to address the gaps in understanding by exploring how age, sex, educational attainment, alcohol consumption, smoking, and health regions influence physical activity patterns in Ontario.
+By examining these relationships, we aim to provide actionable insights that can inform public health policies and interventions tailored to the needs of specific population groups.
+Our findings will be pivotal in developing strategies to increase physical activity participation and improve health equity in Ontario.
+
+The estimand of this paper would be the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on strenuous activities among individuals in Ontario who engage in physical activities.
+
+The remainder of this paper is structured as follows.
+@sec-data would provide a sense of the chosen Computing in the Humanities and Social Sciences dataset as well as data exploration of interested variables.
+@sec-Model goes into the details of the linear regression model fitted on the predictors of interest variables, for which it's results would be showcased in @sec-results.
+Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper.
+At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered.
+
+Overall, this paper was generated using R programming language [@citeR].
+Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow].
+Generation of maps were done using functions in `tmap` package [@citationTmap].
+Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].
 
@@
 This can be found in the data section for Canada[@OtherSource1].
-Also, ParticipACTION provides key statistics regarding proportion of adult population in canada meeting the national guidelines of 150 minutes of moderate to vigorous exercise, as well as how sedentary canadians are [@OtherSource2].
+Also, ParticipACTION provides key statistics regarding proportion of adult population in canada meeting the national guidelines of 150 minutes of moderate to vigorous physical activites, as well as how sedentary canadians are [@OtherSource2].
 However, the availability of raw individual level data is absent from these sources.
@@
 
-Our goal is to explore the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on exercise among individuals in Ontario who engage in regular exercise.
+Our goal is to explore the actual effect of socio-demographic factors, substance abuse, and health factors on how much time is spent on physical activities among active individuals in Ontario.
 The instrument of measurement would be surveys.
-Through the cross-sectional survey conducted by CCHS, we are able to get the number of hours spent on vigorous exercise per individual, our response variable.
+Through the cross-sectional survey conducted by CCHS, we are able to get the number of hours spent on physical activities per individual, our response variable.
 
@@
 Afterwards, they were asked to indicate the total time spent on those activities.
-To address the concern of recall bias where an individual may not recall the specific timing in engaging with exercise, surveyors sectioned vigorous exercise into commuting, sports, work and volunteering.
-Then within each section, surveyors asked which day (Monday to Sunday) in the past week had they engaged in such exercise that made them sweat and breathe a little harder.
+To address the concern of recall bias where an individual may not recall the specific timing in engaging with physical activity, surveyors sectioned physical activity into commuting, sports, work and volunteering.
+Then within each section, surveyors asked which day (Monday to Sunday) in the past week had they engaged in such activities that made them sweat and breathe a little harder.
 By doing so, the activities individuals did during the past week would be anchored on the days of the week.
-Following this would be for the respondents to indicate the total number of hours they engaged in each
-section of vigorous exercise for the past 7 days.
+Following this would be for the respondents to indicate the total number of hours they engaged in each section of physical activities for the past 7 days.
 
@@
 
-For socio-demographic factors possibly affecting time spent on vigorous exercise, the surveyors asked the respondent for their age (DHH_AGE) from 0 years old to 121 years old), sex (DHH_SEX) being either male or female, personal income obtained from tax forms submitted to the Canada Revenue Agency upon consent from respondent, geographical health region (ADM_Q037) that was provided by respondents when they agreed to the phone survey, and highest educational attainment (EHG2_04) ranging from less than high school diploma to master's degree and above.
+For socio-demographic factors possibly affecting time spent on physical activites, the surveyors asked the respondent for their age (DHH_AGE) from 0 years old to 121 years old), sex (DHH_SEX) being either male or female, personal income obtained from tax forms submitted to the Canada Revenue Agency upon consent from respondent, geographical health region (ADM_Q037) that was provided by respondents when they agreed to the phone survey, and highest educational attainment (EHG2_04) ranging from less than high school diploma to master's degree and above.
 
@@
 
-Overall, each entry obtained by respondents assigns quantitative values to each factor and the the response for which we can explore subsequently using linear regression to determine which factors have greater influence in the time spent exercising amongst individuals that are engaged in exercising.
-
-## Response variable: Time spent on vigorous activity in the past 7 days
+Overall, each entry obtained by respondents assigns quantitative values to each factor and the the response for which we can explore subsequently using linear regression to determine which factors have greater influence in the time spent on physical activities amongst active individuals.
+
+## Response variable: Time spent on physical activity in the past 7 days
 
 From the CCHS dataset, we look at paadvmva variable.
-This variable is continuous with values ranging from 0 to 9902 representing the number of minutes spent on vigorous activity in the past 7 days.
-For easier readability, the variable will be renamed as time_spent_vigorous_exercise_7d.
+This variable is continuous with values ranging from 0 to 9902 representing the number of minutes spent on physical activity in the past 7 days.
+For easier readability, the variable will be renamed as $time\_spent\_vigorous\_exercise\_7d$.
 As explained in section @sec-measurement, this variable measures the time spent on any activities lasting more than 10 minutes that makes an individual sweat more than they normally would.
@@
 By using this variable, we are assuming that the routine of an individual from the past 7 days is indicative of their general routine beyond the past 7 days.
-This variable may not account for individuals who only began vigorous sweating routine within the past 7 days and were sedentary before, which may not be helpful if included in subsequent model fitting.
-
-Since we are interested in individuals that are already engaging in strenuous activities, we can filter time_spent_vigorous_exercise_7d to exclude 0.
+This variable may not account for individuals who only began a physical routine involving sweat within the past 7 days and were sedentary before, which may not be helpful if included in subsequent model fitting.
+
+Since we are interested in individuals that are already engaging in strenuous activities, we can filter $time\_spent\_vigorous\_exercise\_7d$ to exclude 0.
 After removing such entries, we obtain the following summary statistics as shown in @tbl-summarised_statistics_time_spent_vigorous_exercise_7d.
-The median time spent on vigorous activities the past 7 days would be 300 minutes, which averages out to 43 minutes per day.
+The median time spent on physical activities the past 7 days would be 300 minutes, which averages out to 43 minutes per day.
 
@@
 #| label: tbl-summarised_statistics_time_spent_vigorous_exercise_7d
-#| tbl-cap: Summarised statistics for time spent on vigorous exercise in the past 7 days
-
-original_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+#| tbl-cap: Summarised statistics for time spent on physical activities in the past 7 days
+
+original_data <- read_parquet("../data/02-analysis_data/analysis_data.parquet")
 
@@
 
-kable(data_cleaned_1 |>
+response_var_summarised_data <- data_cleaned_1 |>
   summarise(`Min.` = min(time_spent_vigorous_exercise_7d),
@@
             `3rd Qu.` = quantile(time_spent_vigorous_exercise_7d, 0.75),
-            `Max.` = max(time_spent_vigorous_exercise_7d)))
+            `Max.` = max(time_spent_vigorous_exercise_7d))
+
+kable(response_var_summarised_data)
 
@@
 
-Interestingly, the maximum time spent on vigorous activities the past 7 days was 9902 minutes, and this averages to 1415 minutes per day, which is equivalent to 23.5 hours in a day.
+Interestingly, the maximum time spent on physical activities the past 7 days was 9902 minutes, and this averages to 1415 minutes per day, which is equivalent to 23.5 hours in a day.
 This is very unlikely as this suggests that an individual is constantly engaging in physical strenuous activity without sleeping.
@@
 Roughly 10% of the observed data are considered outliers as they fall outside 1.5 times the interquartile range from the 25% and 75% quartile values.
-We would be performing log transformation which would be discussed in section (INSERT SECTION HERE PLEASE DONT FORGET) which would address these outlier values.
+We would be performing log transformation on our response variable.
+The motivation for this is detailed in [Appendix -@sec-transform-response].
 
@@
 #| label: fig-boxplot_time_spent_vigorous_exercise_7d
-#| fig-cap: Boxplot of time spent on vigorous exercise in the past 7 days
-#boxplot(data_cleaned_1$time_spent_vigorous_exercise_7d, outcol='indianred4')
-ggplot(data_cleaned_1) +
-  geom_boxplot(aes(y=time_spent_vigorous_exercise_7d), 
-               outlier.color = 'indianred4', 
-               outlier.alpha=0.2) +
+#| fig-cap: Boxplot of time spent on physical activities in the past 7 days
+
+data_cleaned_1 |> 
+  ggplot(aes(y=time_spent_vigorous_exercise_7d, x='')) +
+  geom_boxplot(outlier.color = 'indianred4', 
+               outlier.alpha=0.2,
+               width = 0.5) +
+  geom_jitter(alpha = 0.01, width=0.15) +
   theme_minimal() +
@@
   ylab('') +
-  xlab('Boxplot of time spent on vigorous exercise in the past 7 days')
-```
-
-Subsequently, we would have to remove observations that are very unlikely as they do not represent the general population and could skew the distribution of time spent on vigorous exercise.
+  xlab('Boxplot of time spent on physical activities in the past 7 days')
+```
+
+Subsequently, we would have to remove observations that are very unlikely as they do not represent the general population and could skew the distribution of time spent on physical activities.
 
@@
 #| label: fig-hist-time_spent_vigorous_exercise_7d
-#| fig-cap: Histogram of time spent on vigorous exercise the past 7 days
+#| fig-cap: Histogram of time spent on physical activities the past 7 days
 
@@
      main='',
-     xlab = 'time spent on vigorous exercise the past 7 days / min',
+     xlab = '',
      ylab = 'no. of individuals',
@@
 ```
-## Interested explanatory variables
-
-There are a total of 13 predictors of interest that is seemingly related to the time spent in vigorous exercise an exercising individual gets.
+## Interested explanatory variables {#sec-interested_variables}
+
+There are a total of 13 predictors of interest that is seemingly related to the time spent on physical activities by an individual in Ontario.
 
@@
 #| label: fig-barplot-age
-#| fig-cap: Histogram of time spent on vigorous exercise the past 7 days
+#| fig-cap: Histogram of time spent on physical activites the past 7 days
 
@@
 #| tbl-cap: Summarised statistics of respondent ages
-kable(data_cleaned_1 |>
+
+age_data_summarised_statistics <- data_cleaned_1 |>
   summarise(`Min.` = min(age),
@@
             `3rd Qu.` = quantile(age, 0.75),
-            `Max.` = max(age)))
-```
+            `Max.` = max(age))
+
+kable(age_data_summarised_statistics)
+```
+
 ```{r age predictor boxplot}
@@
 
-ggplot(data_cleaned_1) +
-  geom_boxplot(aes(y=age), 
-               outlier.color = 'indianred4', 
+data_cleaned_1 |>
+  ggplot(aes(y=age, x='')) +
+  geom_boxplot(outlier.color = 'indianred4', 
                outlier.alpha=0.2) +
+  geom_jitter(aes(col = age),
+              alpha=0.1, 
+              width = 0.15) +
   theme_minimal() +
   scale_y_continuous(breaks = c(0:15)) +
+  scale_color_viridis_c(name = 'age bin') +
   theme(axis.text.x = element_blank(),
@@
         panel.grid.minor.x = element_blank()) +
-  ylab('respondent ages') +
+  ylab('respondent age bin') +
   xlab('')
@@
 
-Sex is a possible influential predictor for total time spent on vigorous activity.
+Sex is a possible influential predictor for total time spent on physical activity.
 Historically, men are associated with exercise and involved in more labour intensive roles such as construction which involves heavy lifting.
 According to @womenExerciseWHO, women are less active than men by 5% since 2000.
-However, with the advent of female only establishments such as female gyms, and 32% increase in memberships at fitness and health clubs from 2010 to 2019, it would be interesting to see if the sex of an individual still affects the time spent on vigorous activites among physically active individuals.
+However, with the advent of female only establishments such as female gyms, and 32% increase in memberships at fitness and health clubs from 2010 to 2019, it would be interesting to see if the sex of an individual still affects the time spent on physical activites among physically active individuals.
 
@@
 In 2023, @EducationAndActiveLifestyle has done a study to show that highest educational attainment may be a leading factor for physical activity.
-Therefore, it would be worth exploring the effects of highest levels of education with the time spent on vigorous activities among the Ontario residents.
+Therefore, it would be worth exploring the effects of highest levels of education with the time spent on physical activities among the Ontario residents.
 
@@
 #| tbl-cap: Proportion of respondents grouped by highest educational level
-kable(data_cleaned_1 |>
+
+proportion_highest_educ_table <- data_cleaned_1 |>
   count(highest_educational_attainment, name = 'proportion_of_respondents') |>
@@
            highest_educational_attainment == 3 ~ 'post sec certification and above',
-           .default = '')))
+           .default = ''))
+
+kable(proportion_highest_educ_table)
 
@@
 
-Health region is a categorical variable which tells us the geographical location an individual belongs to defined by Provincial Ministries of Health.
+Health region is a categorical variable which tells us the geographical location an individual belongs to defined by Provincial Ministries of Health (as detailed by documentation provided by CHASS).
 It could be possible that certain health regions have more funding that are projected towards increasing awareness towards having an active lifestyle.
-Therefore, it is worth exploring the relationship between health region and time spent on vigorous activities.
+Therefore, it is worth exploring the relationship between health region and time spent on physical activities.
 
@@
 They are encoded using 5 digits.
-Kindly refer to the appendix (INSERT APPENDIX HERE PLEASE DONT FORGET) to refer to all the various health regions.
 @fig-map_health_region shows us the various health regions in Ontario.
@@
 
-```{r}
+By using a GIS focused library in R called tmap [@citationTmap], we are able to generate the map shown in @fig-map_health_region.
+
+```{r health region map}
 #| echo: false
@@
 #| fig-cap: Map of top five and bottom five occuring Ontario health
-library(tmap)
 
@@
 - 1: No income or income loss
-- 2: Less than $20,000 
+- 2: Less than \$20,000 
 - 3: \$20,000 to \$39,999
@@
 They account for less than 0.5% of the sample.
-In the 2018 annual wages report provided by @PersonalIncomeTaxFiling2018, the annual personal income reported from tax filing in Ontario that year was $39,510.
+In the 2018 annual wages report provided by @PersonalIncomeTaxFiling2018, the annual personal income reported from tax filing in Ontario that year was \$39,510.
 Therefore, the sampling population's personal income data is representative of Ontario individuals.
@@
 
-### Variable 7: Illicit drug use the past 12 months
+### Variable 7: Illicit drug use the past 12 months {#sec-drug_exploration}
 
 This is a binary variable which indicates whether an individual has used drugs in the past 12 months prior to the interview.
-Drugs include marijuana, hashish, cocaine, amphetamines, speed, methamphetamines, crystal meth, ecstasy, hallucinogens, and sniffing glue, gasoline or other solvents.
+Drugs include marijuana, hashish, cocaine, amphetamines, speed, methamphetamine, crystal meth, ecstasy, hallucinogens, and sniffing glue, gasoline or other solvents.
 If they have attempted any of these drugs, this variable would be flagged as '1' and '0' otherwise.
 
-It would be interesting to see if drug consumption has an effect on exercising.
+It would be interesting to see if drug consumption has an effect on physical activeness.
 @PhysicalActivityAndSubstanceAbuse exerts that individuals who are more active tend to have lower substance abuse, suggesting a correlation between these two.
-Having this variable would be interesting to explore the relationship of an individuals drug usage and their time spent on exercise.
+Having this variable would be interesting to explore the relationship of an individuals drug usage and their time spent on physical activities.
 
@@
 
-kable(data_cleaned_1 |>
+drug_use_proportion_table <- data_cleaned_1 |>
   count(illicit_drug_use, name = 'num_of_respondents') |>
@@
                                    'never consumed drugs',
-                                   'consumed at least once')))
+                                   'consumed at least once'))
+
+kable(drug_use_proportion_table)
 ```
@@
 A cross sectional study conducted by @SmokingAndPhysicalActiveness found that smokers are less likely to be physically active compared to non-smokers.
-Hence, it would be interesting to discover if there is an effect of smoking more than 100 cigarettes on the time spent on exercise.
+Hence, it would be interesting to discover if there is an effect of smoking more than 100 cigarettes on the time spent on physical activites.
 
 @fig-barplot-cigarettes shows us that within the sample, the visual difference between the number of respondents who smoked more than 100 cigarettes (blue column) and respondents who smoked lesser than 100 cigarettes are not that big.
-From @tbl-barplot-cigarettes, 57% of respondents have smoked less than 100 cigarettes while the remaining 43% of respondents smoked more than 100 cigarettes.
+Specifically, 57% of respondents have smoked less than 100 cigarettes while the remaining 43% of respondents smoked more than 100 cigarettes.
 As the classes in this variable is almost split half,
@@
 #| label: fig-pyrmaid-cigarettes
-#| tbl-cap: Pyramid of cigarettes consumption in respondent's lifetime by age groups
+#| fig-cap: Pyramid of cigarettes consumption in respondent's lifetime by age groups
 
@@
 
-# Model
+# Model {#sec-Model}
 
 The goal of our model is to explore which factors affect time spent on physical activities. Therefore a simple exploratory linear regression would be the chosen 
-modelling algorithm to determine it.
-A full model where our response (time_spent_on_physical_activities) would be fitted on all our interested variables outlined in @sec-data-interestedvariables (INSERT THE SECTION  HERE LATER PLEASE DONT FORGET).
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+modelling algorithm to determine it. A full model where our response $time\_spent\_vigorous\_exercise\_7d$ would be fitted on all our interested variables outlined in @sec-interested_variables.
 
@@
 \begin{align*} 
-y_k  &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Alc_k +\beta_5I*(Drug_k) + \beta_6I*(Cigar_k) + \\ 
+\ln(y_k)  &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Alc_k +\beta_5I*(Drug_k) + \beta_6I*(Cigar_k) + \\ 
      & \quad\: \beta_7I*(HealthRegion_k=35953) + \beta_8*I(HealthRegion_k=35970) + \epsilon_k
@@
 \begin{align} 
-y_k|\mu_k, \sigma^2 &\sim \mbox{Normal}(\mu_k, \sigma^2) \tag{1} \\
+\ln(y_k)|\mu_k, \sigma^2 &\sim \mbox{Normal}(\mu_k, \sigma^2) \tag{1} \\
 \mu_k &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Alc_k + \tag{2} \\
@@
 
-### Model justification
+### Model justification {#sec-model_justification}
 
@@
 \begin{align*} 
-y_k  &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Income_k +\beta_5Alc_k +\beta_6I*(Drug_k) +  \\ 
+\ln(y_k) &= \beta_0 + \beta_1Age_k + \beta_2Sex_k + \beta_3Educ_k + \beta_4Income_k +\beta_5Alc_k +\beta_6I*(Drug_k) +  \\ 
      & \quad\: \beta_7I*(Cigar_k) + \beta_cI*(HealthRegion=i) + \epsilon_k
@@
 
-# Results
+@tbl-BIC_values shows us the BIC values for both the full model and the reduced model.
+After removing certain explanatory variables, the BIC score has decreased, indicating that the reduced model is has stronger explanatory power while having a less complex model.
+Therefore, the reduced model is chosen from the original full model.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-BIC_values
+#| tbl-cap: Table of BIC values for original and final model
+
+original_model <- 
+  readRDS(file = ("../models/original_model.rds"))
+
+# parse in final model
+final_model <- readRDS('../models/final_model.rds')
+
+BIC_tibble <- as_tibble(
+  list(`Model` = c('Original full model', 'Final reduced model'),
+  `BIC value` = c(BIC(original_model), BIC(final_model))
+  )) |>
+  mutate(`BIC value` = round(`BIC value`, 0))
+
+kable(BIC_tibble)
+```
+
+# Results {#sec-results}
 
@@
 \begin{align*} 
-y_k  &= 5.84 - 0.03Age_k + 0.23 Sex_k - 0.08Educ_k + 0.04Alc_k + 0.15I*(Drug_k) +  \\ 
+\ln(y_k)  &= 5.84 - 0.03Age_k + 0.23 Sex_k - 0.08Educ_k + 0.04Alc_k + 0.15I*(Drug_k) +  \\ 
      & \quad\: 0.12I*(Cigar_k) + \beta_cI*(HealthRegion=i) + \epsilon_k
@@
 
+Overall, the individual t-test for our predictors were found to be below our set threshold of 95% alpha.
+The confidence interval for each predictor is listed in @tbl-confidence_interval_lmmodel shown below.
+
+```{r confidence interval for linear model}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-confidence_interval_lmmodel
+#| tbl-cap: Confidence Interval of linear model
+
+# obtain variable names in lmmodel
+variables <- variable.names(final_model)
+
+# create confidence interval table to display 
+conf_interval_table <- as_tibble(confint(final_model)) |>
+  mutate(`Variable Name` = variables, .before = '2.5 %') |>
+  mutate_if(is.numeric, list(~round(.,2)))
+
+# diaply confidence interval table
+kable(conf_interval_table)
+```
+
 Our results are summarized in @tbl-modelresults.
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-```
-
-```{r}
+```{r summary of linear model}
 #| echo: false
@@
   list(
-    "Time spent on physical activity" = final_model
+    "Log(Time_spent_on_physical_activity)" = final_model
   )
@@
 
-# Discussion
-
-## Males are more likely to engage in physical activities {#sec-men_physical_activites}
+# Discussion {#sec-discussion}
+
+## Males are more likely to engage in physical activities{#sec-men_physical_activites}
+
+According to @fig-physical_activity_by_sex, we can observe that amongst individuals that are active, men spend (on average) more time on physical activity than women.
+In particular, according to our sample men spend on average 639 minutes a week on physical activites while women spend on average 437 minutes a week on physical activities.
+
+```{r bar plot of results - sex}
+#| echo: false
+#| eval: true
+#| label: fig-physical_activity_by_sex
+#| fig-cap: "Bar plot of average time spent on physical activity by sex"
+#| warning: false
+
+options(scipen=999)
+exercise_sex_data <- data_cleaned_1 |>
+  group_by(sex) |>
+  summarise(
+         avg_physical_activity_time_spent = mean(time_spent_vigorous_exercise_7d)) |>
+  mutate(sex = ifelse(sex == 0, 'Female', 'Male')) |>
+  ungroup()
+
+exercise_sex_data|>
+  ggplot() +
+  geom_bar(mapping = aes(x=sex, 
+                         y=avg_physical_activity_time_spent,
+                         fill=sex), 
+           stat = 'identity') +
+  scale_fill_manual(values = c('Female' = 'indianred4',
+                               'Male' = 'skyblue1')) +
+  xlab('') +
+  ylab('Average time spent on physical activity / min') +
+  theme_minimal() +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
 For the explanatory variable `sex`, the corresponding coefficient from the fitted linear model is 0.232.
@@
 
-The linear model is telling us that if an individual has smoked 100 cigarettes in their lifetime, they are 13% more likely to spend more time being physically active than an individual who has not smoked more than 100 cigarettes in the same timeframe.
+The linear model is telling us that if an individual has smoked 100 cigarettes in their lifetime, they are 13% more likely to spend more time being physically active than an individual who has not smoked more than 100 cigarettes in the same time frame.
 While it may seem contradicting to what we believe, there may be merit to this value.
+From @fig-physical_activity_by_smoking_status, we can see that individuals who had smoked 100 cigarettes in their lifetime spent more time on physical activity than individuals who have not smoked 100 cigarettes.
+In particular, our data shows that on average, individuals who smoked 100 cigarettes spend 618 minutes on physical activities in a week while those who don't spend only 418 minute on physical activities.
+
+```{r bar plot of results - smoking status}
+#| echo: false
+#| eval: true
+#| label: fig-physical_activity_by_smoking_status
+#| fig-cap: "Bar plot of average time spent on physical activity by smoking status"
+#| warning: false
+
+cigarette_sex_data <- data_cleaned_1 |>
+  group_by(smoked_hundred_cigarettes) |>
+  summarise(
+         avg_physical_activity_time_spent = mean(time_spent_vigorous_exercise_7d)) |>
+  mutate(smoked_hundred_cigarettes = ifelse(smoked_hundred_cigarettes == 0, 'Non smoker/\nlight smoker', 'Past heavy smoker/\nCurrent heavy smoker'),
+         smoked_hundred_cigarettes = factor(smoked_hundred_cigarettes, levels = c('Non smoker/\nlight smoker', 'Past heavy smoker/\nCurrent heavy smoker'))) |>
+  ungroup()
+
+cigarette_sex_data |>
+  ggplot() +
+  geom_bar(mapping = aes(x=smoked_hundred_cigarettes, 
+                         y=avg_physical_activity_time_spent,
+                         fill=smoked_hundred_cigarettes), 
+           stat = 'identity',
+           show.legend = F) +
+  scale_fill_manual(values = c('Non smoker/\nlight smoker' = 'indianred4',
+                               'Past heavy smoker/\nCurrent heavy smoker' = 'skyblue1')) +
+  xlab('') +
+  ylab('Average time spent on physical activity/ min') +
+  theme_minimal() +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()) 
+```
 
@@
 While smoking may have been affected an individual's physical capabilities, it could have also increased awareness regarding the importance of their health.
-Since most canadians who had a history of smoking had quit smoking, it is very possible that they have greater health conscious.
+Since most Canadians who had a history of smoking had quit smoking, it is very possible that they have greater health conscious.
 Coupled with increased government support for programs with the sole purpose of increasing physical activeness in Ontario [@citationGovernmentSupportExercise], it is no longer surprising why individuals who smoked 100 cigarettes in their lifetime are on average 13% more likely to engage in physical activities.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Possible increase in stimulant use (about drugs)
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
+## Possible increase in stimulant use
+
+According to the model, the coefficient for illicit drug use is a positive value.
+This means that if an individual from Ontario has consumed drugs listed in @sec-drug_exploration, they would spend on average 17% more time on physical activities.
+This is also another surprising finding since we have also discussed in @sec-drug_exploration that there is a negative correlation between drug use and physical activeness, but our findings for our Ontario population suggests a positive correlation.
+From @fig-physical_activity_by_illicit_drug_use, we can see that individuals who consumed drugs listed in @sec-drug_exploration in their lifetime spent more time on physical activity than individuals who have not consumed drugs.
+In particular, our data shows that on average, individuals are drug abusers spend 670 minutes on physical activities in a week while those who don't spend only 503 minute on physical activities.
+
+```{r bar plot of results - illicit drug use}
+#| echo: false
+#| eval: true
+#| label: fig-physical_activity_by_illicit_drug_use
+#| fig-cap: "Bar plot of average time spent on physical activity by illicit drug use"
+#| warning: false
+
+drug_sex_data <- data_cleaned_1 |>
+  group_by(illicit_drug_use) |>
+  summarise(
+         avg_physical_activity_time_spent = mean(time_spent_vigorous_exercise_7d)) |>
+  mutate(illicit_drug_use = ifelse(illicit_drug_use == 0, 'Non-drug users', 'Drug users'),
+         illicit_drug_use = factor(illicit_drug_use, levels = c('Non-drug users', 'Drug users'))) |>
+  ungroup()
+
+drug_sex_data |>
+  ggplot() +
+  geom_bar(mapping = aes(x=illicit_drug_use, 
+                         y=avg_physical_activity_time_spent,
+                         fill=illicit_drug_use), 
+           stat = 'identity',
+           show.legend = F) +
+  scale_fill_manual(values = c('Non-drug users' = 'indianred4',
+                               'Drug users' = 'skyblue1')) +
+  xlab('') +
+  ylab('Average time spent on physical activity/ min') +
+  theme_minimal() +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank()) 
+```
+
+Of the listed drugs, amphetamine and methamphetamine were identified as well.
+According to @citationDrugsAndActiveLifestyle, they mentioned that the use of meth (short term for methamphetamine) leads to increased concentrations of dopamine, which is a mood enhancer.
+This is coupled with the increase of body movement.
+Likewise, in a study conducted by the Indiana University [@citationAmphetamineandExercise], they found that the use of Amphetamine masks fatigue, allowing for individuals to exercise longer.
+This shows that it is possible that the inclusion of these drugs could have explained the surprising increase in time spent on physical activities despite the expectation that an individual who abuses drugs would have lower exercise.
+
+The inclusion of these drugs in the survey could explain the positive coefficient seen from our model.
+However, this is just a hypothesis and an experiment would need to be conducted to explore if this finding is true, where we obtain a sample of drug users who consumed amphetamine and methamphetamine and drug users who did not consume the two drugs.
+
+## Weaknesses
+
+Self-reported bias is a common limitation in health data collection, particularly when examining behaviors like physical activity, alcohol consumption, and smoking.
+Participants may overestimate or underestimate their activities due to memory recall issues, social desirability bias, or personal perceptions of what constitutes "healthy" behavior.
+For example, individuals might inflate their reported exercise levels or downplay unhealthy habits like smoking.
+Such inaccuracies can introduce systematic error, leading to biased estimates of associations and potentially undermining the validity of the study's conclusions.
+Mitigating this requires complementary objective measures, such as wearable activity trackers or biomarkers.
 \newpage
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
+# Diagnostic Checks {#sec-diagnostic_checks}
+## Motivation to transform response variable for linear regression model {#sec-transform-response}
+
+Our response variable was log transformed to fulfill diagnostic checks for linear regression model.
+Observe the change in qq plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log transformed.
+From the left plot of @fig-diagnostic_plot_qq, we can see that values of standardised residuals are a lot larger than expected at higher quantile values.
+This suggests that the distribution of the residuals is severely right skewed.
+This contradicts the assumption of normality of residuals, which would make our linear regression model unreliable.
+
+From the right plot of @fig-diagnostic_plot_qq, after we transform our response using the natural logarithm function ($ln$), we can see that most of the residuals lie on the y=x line in the qqplot, suggesting that the sampling distribution is approximately normal.
+
+```{r qq plot}
+#| echo: false
+#| eval: true
+#| label: fig-diagnostic_plot_qq
+#| fig-cap: "Diagnostic plot: QQ Plot"
+#| warning: false
+
+not_transformed_model <-
+  readRDS(file = ("../models/original_model_not_transformed.rds"))
+
+par(mfrow = c(1,2))
+plot(not_transformed_model,2, col=adjustcolor('black', alpha=0.2), sub='')
+title(main='not transformed')
+p2_1 <- plot(original_model,2, col=adjustcolor('black', alpha=0.2), sub='')
+title(main='log transformed')
+```
+
+Now, we observe the change in the scale location plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log transformed.
+From the left plot of @fig-diagnostic_plot_scale_location, we can see a weak but visible trend that as the fitted values increase, so does the square root of the standardized residuals.
+This suggests that the residuals are not randomly distributed, which goes against the assumption of linear regression that residuals are independent from each other.
+Furthermore, the mean red line is not horizontally flat.
+This is suggesting a positive trend.
+This tells us that the variance of the residuals is increasing, therefore implying that our residuals do not satisfy homoscedasticity assumption of linear regression, which would lead to an unreliable linear regression model.
+
+From the right plot of @fig-diagnostic_plot_scale_location, we can see that the standardised residuals are randomly scattered across the plot across all fitted values.
+This suggests that residuals are independant from each other.
+Furthermore, the mean red line is (almost) horizontally flat, for which suggests that the variance of residuals across all fitted values are constant.
+Therefore, homoscedasticity assumption for linear regression is also satisfied, allowing us to interpret the coefficient estimates of our fitted linear model with greater reliability.
+
+```{r scale location plot}
+#| echo: false
+#| eval: true
+#| label: fig-diagnostic_plot_scale_location
+#| fig-cap: "Diagnostic plot: Scale location"
+#| warning: false
+
+par(mfrow = c(1,2))
+plot(not_transformed_model,3, sub='')
+title(main='not transformed')
+plot(original_model,3, sub='')
+title(main='log transformed')
+```
+
+By performing log transformation of my response, we have satisfied the linear regression assumptions, allowing us to interpret from a reliable linear regression model, therefore explaining the need to transform our response variable in @sec-Model.
+This would also allow us to perform model reduction outlined in @sec-model_justification
+
+## Diagnostic check for final linear model after BIC
+
+The final linear model was obtained after performing backward feature elimination using BIC score as the criterion.
+In this section, we will be showing that the final model satisfies linear regression assumptions.
+
+@fig-diagnostic_plot_qq_final_model shows us that the values of standardised residuals are approximately equal to the theoretical quantiles, which suggests that the residuals are normally distributed.
 ```{r}
+#| echo: false
 #| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
+#| label: fig-diagnostic_plot_qq_final_model
+#| fig-cap: "Diagnostic plot: QQ Plot"
+#| warning: false
+plot(final_model,2, sub='')
+```
+From @fig-diagnostic_plot_scale_location_final_model, we can see that the standardised residuals are randomly scattered across the scale location plot, which suggests that residuals are independant from each other.
+Furthermore the mean red line is (almost) horizontal which suggests that variance of residuals are constant across all fitted values.
+This means that the residuals obtained from the final model satisfies homoscedasticity assumption in linear regression
 ```{r}
@@
 #| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-```
+#| label: fig-diagnostic_plot_scale_location_final_model
+#| fig-cap: "Diagnostic plot: Scale location"
+#| warning: false
+plot(final_model,3, sub='')
+```
+
+Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into relationship between our predictors and the response as discussed in @sec-discussion.
 

### 2024-12-02 scripts/03-clean_data.R

@@ -9,6 +9,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow) # for parquet file saving
 
 #### Download data ####
 raw_data <- read_csv("data/01-raw_data/raw_data.csv", na='.')
@@ -152,6 +153,8 @@ data_intermediate3 <- data_intermediate2 |>
          )
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(data_intermediate3, "data/02-analysis_data/analysis_data.csv")
+# save analysis data as a parquet file
+write_parquet(
+  x=data_intermediate3,
+  sink="data/02-analysis_data/analysis_data.parquet"
+)

### 2024-12-02 scripts/06-model_data.R

@@ -10,17 +10,26 @@
 #### Workspace setup ####
 library(tidyverse)
 library(fastDummies)
+library(arrow) # reading parquet file
+
 
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv") |>
-  filter(time_spent_vigorous_exercise_7d != 0)  |>
-  mutate(#health_region = factor(health_region),
-         time_spent_vigorous_exercise_7d = log(time_spent_vigorous_exercise_7d)) |>
+analysis_data_not_transformed <- read_parquet("data/02-analysis_data/analysis_data.parquet") |>
+  filter(time_spent_vigorous_exercise_7d != 0) |>
   select(-province)
+
+analysis_data <-  analysis_data_not_transformed |>
+  mutate(#health_region = factor(health_region),
+         time_spent_vigorous_exercise_7d = log(time_spent_vigorous_exercise_7d)) 
 
 # run this if generated column names are in scientific notation
 # options(scipen=0)
 analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
+  select(-c(health_region, has_mood_disorders, 
+            perceived_life_stress,
+            perceived_work_stress))
+
+analysis_data_not_transformed <- dummy_cols(analysis_data_not_transformed, select_columns='health_region', remove_first_dummy = TRUE) |>
   select(-c(health_region, has_mood_disorders, 
             perceived_life_stress,
             perceived_work_stress))
@@ -52,6 +61,19 @@ original_lmmodel <- lm(time_spent_vigorous_exercise_7d ~ ., #+
                           # num_alc_drank_12m,
               data = analysis_data_w_dummies)
 
+original_lmmodel_not_transformed <- lm(time_spent_vigorous_exercise_7d ~ ., #+
+                       # age:highest_educational_attainment +
+                       # age:sex +
+                       # sex:personal_income,
+                       # age+ 
+                       # highest_educational_attainment+
+                       # personal_income+
+                       # smoked_hundred_cigarettes+
+                       # # perceived_work_stress+
+                       # # illicit_drug_use+
+                       # health_region +
+                       # num_alc_drank_12m,
+                       data = analysis_data_not_transformed)
 # obtain summary plots to check linear regression assumptions
 # 1. normality of sampling distribution
 # 2. homoscedasticity (random variance)
[…]

### 2024-12-03 paper/paper.qmd
Final Draft Edits
1. Minor formatting edits in paper.qmd

2. Addition of survey exploration in second section of appendix in paper.qmd

3. Edited preamble of all scripts

4. Added simulation script and testing scripts for both simulated data and actual data

5. Added LLM interaction and updated readME to reflect added interaction

6. Removed unnecessary files

@@
 ---
-title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers"
+title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers respectively"
 subtitle: "Women are engaging in lesser physical activities than men.
@@
   - Chris Yong Hong Sen
-thanks: "Code and data are available at: [https://github.com/Monoji77/Alcohol_Use_Ontario](https://github.com/Monoji77/Alcohol_Use_Ontario)."
+thanks: "Code and data are available at: [https://github.com/Monoji77/physical_health_ontario](https://github.com/Monoji77/physical_health_ontario)."
 date: today
@@
 Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels.
-These insights highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing valuable information for designing targeted public health interventions to promote physical activity."
+These findings highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing helpful information for designing targeted public health interventions to promote physical activity."
 format: 
@@
 This is especially concerning as physical inactivity is a leading contributor to chronic diseases such as heart disease, diabetes, and mental health disorders.
-Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play a critical role in shaping physical activity behaviors.
-Understanding these factors and their complex interactions is crucial for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.
+Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play an important role in shaping physical activity behaviors.
+Understanding these factors and their complex interactions is important for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.
 
@@
 Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper.
-At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered.
+At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered, and in [Appendix -@sec-survey_methodology], it contains a section regarding the survey and it's methodology.
 
@@
 Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow].
-Generation of maps were done using functions in `tmap` package [@citationTmap].
+Shape date was read into R using `sf` package, and generation of maps were done using functions in `tmap` package [@citeSF; @citationTmap].
 Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].
@@
 The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada.
-The goal of the CCHS survey is to collect comprehensive and reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
+The goal of the CCHS survey is to collect reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
 It aims to provide timely, accessible, and flexible data for monitoring health trends, studying small populations and rare characteristics, and addressing emerging health issues, ultimately helping improve the health and well-being of Canadians.
@@
 According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents).
-Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected ing @fig-barplot-sex.
+Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected in @fig-barplot-sex.
 ```{r sex predictor barplot}
@@
 
-source("../scripts/08-health_region_map.R")
+source("../scripts/05-health_region_map.R")
 
@@
 
-Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into relationship between our predictors and the response as discussed in @sec-discussion.
+Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain understanding in the relationship between our predictors and the response as discussed in @sec-discussion.
+
+# Survey: Canadian Community Health Survey (CCHS) Annual component {#sec-survey_methodology}
 
@@
 
+## CCHS Background
+
+The purpose of CCHS is to provide cross sectional estimates of factors related to different health determinants in Canada.
+The use case for the Canadian health dataset is non-exhaustive and many have used it for the betterment of Canada.
+In the department of sociology and social work, University of Toronto, researchers have used the weight and weight perception data from the 2017-2018 CCHS to predict adolescents mental health, and they were able to find out that normal weighted girls are not immune from poor mental health arising from poor weight perceptions, hence urging specialists in the mental health field to increase support for these girls that have not been receiving enough avenues [@citationUofTCCHS].
+The Canadian Partnership Against Cancer also uses CCHS data to access the severity of the cervical cancer among Canadians, streamlining their cancer screening process as a result and urging the country to focus resources to address inequalities in access to cervical cancer prevention and treatment options between indigenous communities and poor regions [@citationCanadianPartnership2; @citationCanadianPartnership1].
+The one thing that ties many of the health-related studies arising from the use of CCHS studies is to address underrepresented communities in Canada.
+
+With the purpose of unifying data across health sources, the Statistics of Canada had began conducting surveys yearly since 2007.
+From 2015, there was a redesign to their survey methodology that is used for the dataset that was used in this paper.
+
+## CCHS Target Population
+
+Their target population covers individuals aged 12 and above living in all ten provinces of Canada. The target population excludes the following...
+
+- individuals living on reserves, more commonly known as indigenous populations with their own governance within Canada ()
+- full time servicemen in the Canadian Armed Forces
+- youths aged 12 to 17 residing in foster homes
+- individuals residing in jail, nursing homes, mental health institutions and shelters
+- individuals in quebec health region of Nunavik and Terres-Cries-de-la-Baie-James
+
+According to CCHS, they account for lesser than 3% of the population
+
+## CCHS Sampling Frame
+
+The sampling frame for the Canadian Community Health Survey (CCHS) consists of two distinct components based on the target population.
+For the adult population (18 years and older), the sampling frame is derived from an area frame used by the Canadian Labour Force Survey (LFS).
+This area frame includes all primary sampling units (PSUs), which are geographic clusters of dwellings within health regions across Canada.
+For the youth population (12 to 17 years old), the sampling frame is based on a list derived from the Canadian Child Benefit (CCB) files, which includes a list of all program beneficiaries with their names, addresses, and phone numbers.
+
+## CCHS Sample
+The sample for the CCHS is drawn from the respective sampling frames for adults and youth.
+Approximately 130,000 individuals are selected for the survey, with around 120,000 respondents allocated to the adult population and 10,000 to the youth population.
+The sample is stratified by health region, ensuring that each health region has a minimum of 500 respondents.
+The sample for adults is selected from the area frame using a multi-stage sampling process, while the youth sample is directly selected from the CCB list.
+The allocation for provinces and territories is based on population size, with a focus on achieving representativeness within each health region.
+
+## CCHS Data Collection
+Sample Recruitment for the 2018 Canadian Community Health Survey (CCHS)
+The sample for the 2018 CCHS was recruited through a multi-stage, stratified process designed to obtain a representative sample of the Canadian population, aged 12 years and older.
+The sampling process was divided into two main categories based on the target age groups: individuals aged 18 and older and youths aged 12 to 17.
+
+\noindent
+The sample was drawn from two distinct frames:
+
+- Area Frame: This frame was used to select households for respondents aged 18 and older.
+The area frame covers the entire Canadian population and is designed to represent geographic areas across the country.
+- Canadian Child Benefit (CCB) Frame: This frame was used to select households with youths aged 12 to 17.
+The CCB frame was specifically used to identify youth in households receiving the Canadian Child Benefit.
+
+\noindent
+The data collection for the CCHS was carried out using two primary methods:
+
+- Computer-Assisted Telephone Interviewing (CATI): This method was used for approximately 75% of respondents selected from the area frame and for all respondents selected from the CCB frame (aged 12-17).
+Interviews were conducted over the phone by trained interviewers in centralized call centers.
+- Computer-Assisted Personal Interviewing (CAPI): This method was used for about 25% of the area frame respondents.
+These interviews were conducted face-to-face by decentralized field interviewers using laptops.
+Interviewers primarily visited households in person, though telephone follow-ups were permitted if necessary.
+
+\noindent
+The sampling process was designed to maximize coverage and minimize non-response.
+Each regional office of Statistics Canada was responsible for assigning cases to interviewers, with careful consideration of the geographic and demographic distribution of respondents.
+The sample was stratified by health region to ensure representation from each geographic area.
+Interviewers were assigned cases based on their regional collection office, and the assignments typically included no more than 15 cases per interviewer to ensure manageable workloads.
+These assignments were planned to be completed in four non-overlapping three-month collection periods, with different phases for initial and follow-up attempts.
+
+\noindent
+To optimize the contact process, the interviewers first initiated contact with the selected households by either telephone or personal visit, depending on the sampling frame and mode of data collection.
+If initial contact was unsuccessful, additional efforts were made to reach the household through scheduled callbacks, including multiple attempts at different times and days.
+If necessary, personal visits were conducted to obtain the interview.
+
+\noindent
+For youths aged 12-17 selected for interviews, interviewers were required to obtain verbal consent from a parent or guardian before interviewing youths aged 12-14.
+If a parent or guardian requested to see the questionnaire, they were provided with a copy before the interview.
+In cases where the selected youth was unable to complete the interview, a "Person Most Knowledgeable" (PMK) within the household was interviewed instead.
+This procedure ensured that household-level information was collected even if the selected youth was unable to provide it.
+
+\noindent
+If the selected respondent was unable to participate due to physical or mental health conditions, a proxy interview was conducted.
+A knowledgeable household member would provide information on behalf of the selected respondent.
+However, the more personal and sensitive questions were omitted from these interviews, as the proxy could not reliably answer them.
+
+\noindent
+Throughout the recruitment process, several quality control measures were in place to ensure the accuracy and validity of the data.
+A random selection of interviews was validated to ensure the integrity of the data.
+These validations involved contacting households to confirm that the interview took place and checking the quality of the responses.
+Cumulative reports were generated at the end of each collection period, showing response rates by region, to help regional offices identify areas that required additional follow-up.
+Specific sections of interviews were recorded for quality control purposes and analyzed to ensure proper procedures were followed by interviewers.
+
+## CCHS Nonresponse and Nonresponse handling
+
+The Canadian Community Health Survey (CCHS) employs a systematic approach to address nonresponse at both the household and person levels.
+This ensures that the survey results remain representative of the general population, even when certain individuals or households do not respond.
+The nonresponse handling methodology is as follows:
+
+\noindent
+1. Household Nonresponse \newline
+
+Nonresponse at the household level occurs when a sampled household either refuses to participate, is unreachable, or provides unusable data.
+To adjust for nonresponse, the survey sample is divided into response homogeneity groups (RHGs), which group households based on their likelihood of responding.
+This helps ensure that the adjustment for nonresponse is made based on similar response propensities.
+A logistic regression model is used to predict the likelihood of a household responding to the survey.
+Based on the predicted response probabilities, an adjustment factor is computed within each RHG.
+This factor helps correct for nonresponse by redistributing the weight of nonresponding households to those that did respond.
+The final adjustment is made by multiplying the weight of responding households by the adjustment factor, resulting in a new weight.
+Nonresponding households are excluded from the weighting process after this adjustment.
+
+\noindent
+2. Person-Level Nonresponse \newline
+
+The interviewer first collects a complete roster of household members.
+A selected individual from the roster is then interviewed.
+In cases where the selected individual cannot be contacted or refuses to participate, person-level nonresponse occurs.
+Similar to household nonresponse, an adjustment is applied to the weights of responding individuals to account for those who did not respond.
+This ensures that the survey remains representative of the entire population.
+A logistic regression model is applied to estimate the likelihood that a selected individual will respond to the survey.
+This model uses information gathered during the first part of the interview (e.g., age, sex, etc.) and the location of the household.
+The selected individuals are then grouped into response homogeneity groups (RHGs) based on their calculated response probabilities.
+Within each RHG, an adjustment factor is calculated to correct for person-level.
+
+By including factors such as geographic location and household characteristics in the nonresponse adjustment models, CCHS enhances the likelihood that the adjusted sample is representative of diverse populations across different regions and demographic groups.
+Also, the distinct handling of household nonresponse (via weight adjustment A6) and person nonresponse (via weight adjustment A9) ensures that both levels of nonresponse are addressed separately, which helps maintain the accuracy of estimates at both the household and individual levels.
+
+However, the use of multiple weighting adjustments, while intended to improve representativeness, can sometimes lead to overfitting or over-adjustment.
+This may result in a dataset that over-represents certain characteristics at the expense of others, especially if the logistic regression models are not perfectly specified or if important variables are omitted.
+Furthermore, even with sophisticated adjustments, response bias — such as social desirability bias or recall bias — can still affect the dataset.
+Respondents may provide inaccurate or socially desirable answers, and the weighting adjustments cannot fully correct for these biases, especially if they are not uniform across all demographic groups.
+
+\newpage
+
 # References

### 2024-12-03 scripts/00-simulate_data.R

@@ -1,52 +1,54 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of CCHS Health Data in Ontario
+# Author: Chris Yong Hong Sen 
+# Date: 02 December 2024
+# Contact: luke.yong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# Pre-requisites: 
+#   - The `tidyverse` package must be installed and loaded
 
 #### Workspace setup ####
 library(tidyverse)
 set.seed(853)
 
-
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+
+# Create a dataset by randomly assigning values from response variable and explanatory variables
+simulated_data <- tibble(
+  
+  ### response ###
+  # time spent on physical activity
+  time_spent_physical_activity_7d = sample(x=seq(0, 10080, 1), 
+                                           size=200, 
+                                           replace=T),
+  ### predictors ###
+  
+  # frequency of alcohol drank 
+  freq_alc_drank = sample(seq(1,7,1), 200,replace=T),
+  
+  # sex 
+  sex = sample(seq(0,1), 200, replace=T),
+  
+  # age
+  age = sample(seq(3,14), 200, replace=T),
+  
+  # personal income bin
+  personal_income = sample(1:6, 200, replace = T, prob = c(0.01, 0.3, 0.4,0.2,0.2, 0.05)),
+  
+  # highest educational attainment in bins
+  highest_educational_attainment = factor(sample(seq(1,3,1), 200, replace = T, prob=c(0.01, 0.4,0.5))),
+  
+  # health regions in ontario (34 of them)
+  health_region = sample(seq(1,34,1), 200, replace=T),
+  
+  # smoked 100 cigarettes in lifetime
+  smoking_status = sample(0:1, 200, replace=T, prob=c(0.7, 0.3)),
+  
+  # illicit drug use in one's lifetime
[…]

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -1,14 +1,13 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated CCHS dataset
+# Author: Chris Yong Hong Sen 
+# Date: 02 December 2024
+# Contact: luke.yong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? Make sure you are in the `physical_activity_ontario` rproj
 
 
 #### Workspace setup ####
@@ -26,45 +25,87 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 200 rows
+if (nrow(analysis_data) == 200) {
+  message("Test Passed: The dataset has 200 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 200 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 9 columns
+if (ncol(analysis_data) == 9) {
+  message("Test Passed: The dataset has 9 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 9 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
+# Check if all values in the 'time_spent_physical_activity_7d' response column
+# are within valid time range of 7 weeks in minutes
+if (all(between(analysis_data$time_spent_physical_activity_7d, 0, 10080))) {
+  message("Test Passed: All values in response column are valid.")
 } else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
+  stop("Test Failed: The response column contains invalid values.")
 }
 
-# Check if the 'state' column contains only valid Australian state names
[…]

### 2024-12-03 scripts/02-clean_data.R

@@ -1,0 +1,165 @@
+#### Preamble ####
+# Purpose: clean data obotained from https://sda-artsci-utoronto-ca.myaccess.library.utoronto.ca/sdaweb/analysis/?dataset=cchs2017
+# Author: Chris Yong Hong Sen
+# Date: 20 November 2024
+# Contact: luke.yong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - follow steps on github readME to download raw dataset 
+# - knowledge about parsing csv files using tidyverse functinos
+# Any other information needed? Make sure you are in the `physical_activity_ontario` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow) # for parquet file saving
+
+#### Download data ####
+raw_data <- read_csv("data/01-raw_data/raw_data.csv", na='.')
+
+#### Clean data ####
+
+# perform simple col name change and ontario filtering
+data_intermediate_1 <- raw_data |>
+  janitor::clean_names() |>
+  
+  # rename variables for easier interprebility and variable handling
+  rename(
+    # response
+    time_spent_vigorous_exercise_7d = paadvmva,
+    
+    # potential predictors
+    num_alc_drank_12m  = alc_015,
+    bmi_class = hwtdgbcc,
+    has_mood_disorders = ccc_195,
+    age = dhhgage,
+    sex = dhh_sex,
+    illicit_drug_use = drgdvya,
+    highest_educational_attainment = ehg2dvr3,
+    perceived_life_stress = gen_020,
+    perceived_work_stress = gen_025,
+    health_region = geodghr4,
+    personal_income = incdgper,
+    smoked_hundred_cigarettes = smk_020,
+    
+    # variable to filter province to ontario
+    province = geo_prv
+  ) |>
+  
+  
+  filter(
+    
+    # filter for ontario
+    province == 35, 
+    
+    # filter for individuals who are above the age of 18
+    age > 2) 
+
+
+
+
+##############################
+# cleaning description: perform missing value handling for each variable
+# 
+# NA-inclusive columns:
+#   num_alc_drank_12m; has_mood_disorders; illicit_drug_use; highest_educational_attainment; 
+#   perceived_life_stress; perceived_work_stress; bmi_class; personal_income; 
[…]

### 2024-12-03 scripts/03-clean_data.R

@@ -1,160 +1,0 @@
-#### Preamble ####
-# Purpose: clean data obotained from https://sda-artsci-utoronto-ca.myaccess.library.utoronto.ca/sdaweb/analysis/?dataset=cchs2017
-# Author: Chris Yong Hong Sen
-# Date: 20 November 2024
-# Contact: luke.yong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: knowledge about parsing csv files using tidyverse functinos
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow) # for parquet file saving
-
-#### Download data ####
-raw_data <- read_csv("data/01-raw_data/raw_data.csv", na='.')
-
-#### Clean data ####
-
-# perform simple col name change and ontario filtering
-data_intermediate_1 <- raw_data |>
-  janitor::clean_names() |>
-  
-  # rename variables for easier interprebility and variable handling
-  rename(
-    # response
-    time_spent_vigorous_exercise_7d = paadvmva,
-    
-    # potential predictors
-    num_alc_drank_12m  = alc_015,
-    bmi_class = hwtdgbcc,
-    has_mood_disorders = ccc_195,
-    age = dhhgage,
-    sex = dhh_sex,
-    illicit_drug_use = drgdvya,
-    highest_educational_attainment = ehg2dvr3,
-    perceived_life_stress = gen_020,
-    perceived_work_stress = gen_025,
-    health_region = geodghr4,
-    personal_income = incdgper,
-    smoked_hundred_cigarettes = smk_020,
-    
-    # variable to filter province to ontario
-    province = geo_prv
-  ) |>
-  
-  
-  filter(
-    
-    # filter for ontario
-    province == 35, 
-    
-    # filter for individuals who are above the age of 18
-    age > 2) 
-
-
-
-
-##############################
-# cleaning description: perform missing value handling for each variable
-# 
-# NA-inclusive columns:
-#   num_alc_drank_12m; has_mood_disorders; illicit_drug_use; highest_educational_attainment; 
-#   perceived_life_stress; perceived_work_stress; bmi_class; personal_income; 
-#   time_spent_vigorous_exercise_7d; smoked_hundred_cigarettes;
-#
[…]

### 2024-12-03 scripts/03-test_analysis_data.R

@@ -1,0 +1,100 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the final cleaned CCHS dataset
+# Author: Chris Yong Hong Sen 
+# Date: 02 December 2024
+# Contact: luke.yong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 02-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `physical_activity_ontario` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+
+nrow(analysis_data)
+#### Test data ####
+# Test that the dataset has 14,121 rows - there are 14,121 Ontario individuals in 
+# the CCHS dataset 
+test_that("dataset has 14,121 observations", {
+  expect_equal(nrow(analysis_data), 14121)
+})
+
+# Test that the dataset has 9 columns
+test_that("dataset has 9 columns", {
+  expect_equal(ncol(analysis_data), 9)
+})
+
+# Test that the all columns is numeric type
+test_that("all columns are numeric", {
+  numeric_check <- sapply(analysis_data, is.numeric)
+  expect_true(all(numeric_check), "not all columns are numeric")
+})
+
+# Test that there are no missing values in the dataset
+test_that("no missing values in dataset", {
+  expect_true(all(!is.na(analysis_data)))
+})
+
+# Test that there are no missing values in the dataset
+test_that("no missing values in dataset", {
+  expect_true(all(!is.na(analysis_data)))
+})
+
+
+# Test if all values in the 'time_spent_physical_activity_7d' response column
+# are within valid time range of 7 weeks in minutes
+test_that("All values in response column are valid.", {
+  expect_true(all(between(analysis_data$time_spent_vigorous_exercise_7d, 0, 10080)))
+})
+
+# Check if all values in the 'freq_alc_drank' column have valid bins
+# from 1 to 7 inclusive
+test_that("All values in freq_alc_drank column are valid.", {
+  expect_true(all(between(analysis_data$num_alc_drank_12m, 1, 7)))
+})
+
+# Check if all values in the 'sex' column are strictly 0 (females) and 1 (males)
+test_that("'sex' column has valid observations 0 or 1.", {
+  expect_true(all(analysis_data$sex %in% c(0,1)))
+})
+
+# Check if all values in the 'age' column have correct bins from  0 (females) 
[…]

### 2024-12-03 scripts/04-model_data.R

@@ -1,0 +1,197 @@
+#### Preamble ####
+# Purpose: Simple Linear Regression Model
+# Author: Chris Yong Hong Sen 
+# Date: 24 November 2024
+# Contact:luke.yong@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 03-clean_data.R must have been run
+# - Knowledge about linear regression models, assumptions about sampling distribution
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(fastDummies)
+library(arrow) # reading parquet file
+
+
+#### Read data ####
+analysis_data_not_transformed <- read_parquet("data/02-analysis_data/analysis_data.parquet") |>
+  filter(time_spent_vigorous_exercise_7d != 0) 
+
+analysis_data <-  analysis_data_not_transformed |>
+  mutate(#health_region = factor(health_region),
+         time_spent_vigorous_exercise_7d = log(time_spent_vigorous_exercise_7d)) 
+
+# run this if generated column names are in scientific notation
+# options(scipen=0)
+analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
+  select(-health_region)
+
+analysis_data_not_transformed <- dummy_cols(analysis_data_not_transformed, select_columns='health_region', remove_first_dummy = TRUE) |>
+  select(-health_region)
+
+colnames(analysis_data_w_dummies)
+### Initial linear regression model ####
+#
+# predictors: all + interaction terms
+# interaction term: 
+#    1. age:highest_educational_attainment
+#    2. age:sex
+#    3. sex:personal_income
+colnames(analysis_data_w_dummies)
+
+analysis_data_w_dummies |>
+  count(personal_income)
+
+original_lmmodel <- lm(time_spent_vigorous_exercise_7d ~ ., #+
+              # age:highest_educational_attainment +
+              # age:sex +
+              # sex:personal_income,
+                          # age+ 
+                          # highest_educational_attainment+
+                          # personal_income+
+                          # smoked_hundred_cigarettes+
+                          # # perceived_work_stress+
+                          # # illicit_drug_use+
+                          # health_region +
+                          # num_alc_drank_12m,
+              data = analysis_data_w_dummies)
+
+original_lmmodel_not_transformed <- lm(time_spent_vigorous_exercise_7d ~ ., #+
+                       # age:highest_educational_attainment +
+                       # age:sex +
[…]

### 2024-12-03 scripts/05-health_region_map.R

@@ -1,0 +1,76 @@
+#### Preamble ####
+# Purpose: Map out health regions from daataset
+# Author: Chris Yong Hong Sen 
+# Date: 02 December 2024
+# Contact:luke.yong@mail.utoronto.ca 
+# License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - The `tmap` and `sf` package must be installed and loaded
+# - 03-clean_data.R must have been run
+# - Knowledge about linear regression models, assumptions about sampling distribution
+
+library(sf)
+library(tmap)
+library(tidyverse)
+health_region_shp <- st_read(
+  "../data/03-map_data/Health Regions Ontario Shapes/HR_035a18a_e.shp",
+  quiet = T)
+
+# tm_shape(canada_shp) +
+#   tm_borders() +
+tmap_mode('plot')
+
+HR_UID_top <- tibble(HU_ID = top_5_regions) |>
+  mutate(health_region = case_when(
+    HU_ID == 35951 ~ 'City of Ottawa Health Unit',
+    HU_ID == 35953 ~ 'Peel Regional Health Unit',
+    HU_ID == 35960 ~ 'Simcoe Muskoka District Health Unit',
+    HU_ID == 35970 ~ 'York Regional Health Unit',
+    HU_ID == 35995 ~ 'City of Toronto Health Unit'
+  ))
+
+HR_UID_btm <- tibble(HU_ID = bottom_5_regions) |>
+  mutate(health_region = case_when(
+    HU_ID == 35931 ~ 'Elgin-St Thomas Health Unit',
+    HU_ID == 35935 ~ 'Haliburton, Kawartha, Pine Ridge District ',
+    HU_ID == 35940 ~ 'Chatham-Kent Health Unit',
+    HU_ID == 35942 ~ 'Lambton Health Unit',
+    HU_ID == 35956 ~ 'Porcupine Health Unit'
+  ))
+
+
+
+HR_UID_top <- paste0(substr(top_5_regions, 1, 2), substr(top_5_regions, 4, 5))
+
+HR_UID_btm <- paste0(substr(bottom_5_regions, 1, 2), substr(bottom_5_regions, 4, 5))
+
+top_health_regions_shp <- health_region_shp |>
+  filter(HR_UID %in% HR_UID_top)
+
+btm_health_regions_shp <- health_region_shp |>
+  filter(HR_UID %in% HR_UID_btm)
+
+
+health_region_map <- tm_basemap('CartoDB.Positron') +
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,200 +1,0 @@
-#### Preamble ####
-# Purpose: Simple Linear Regression Model
-# Author: Chris Yong Hong Sen 
-# Date: 24 November 2024
-# Contact:luke.yong@mail.utoronto.ca 
-# License: MIT
-# Pre-requisites: Knowledge about linear regression models, assumptions about sampling distributino
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(fastDummies)
-library(arrow) # reading parquet file
-
-
-#### Read data ####
-analysis_data_not_transformed <- read_parquet("data/02-analysis_data/analysis_data.parquet") |>
-  filter(time_spent_vigorous_exercise_7d != 0) |>
-  select(-province)
-
-analysis_data <-  analysis_data_not_transformed |>
-  mutate(#health_region = factor(health_region),
-         time_spent_vigorous_exercise_7d = log(time_spent_vigorous_exercise_7d)) 
-
-# run this if generated column names are in scientific notation
-# options(scipen=0)
-analysis_data_w_dummies <- dummy_cols(analysis_data, select_columns='health_region', remove_first_dummy = TRUE) |>
-  select(-c(health_region, has_mood_disorders, 
-            perceived_life_stress,
-            perceived_work_stress))
-
-analysis_data_not_transformed <- dummy_cols(analysis_data_not_transformed, select_columns='health_region', remove_first_dummy = TRUE) |>
-  select(-c(health_region, has_mood_disorders, 
-            perceived_life_stress,
-            perceived_work_stress))
-
-colnames(analysis_data_w_dummies)
-### Initial linear regression model ####
-#
-# predictors: all + interaction terms
-# interaction term: 
-#    1. age:highest_educational_attainment
-#    2. age:sex
-#    3. sex:personal_income
-colnames(analysis_data_w_dummies)
-
-analysis_data_w_dummies |>
-  count(personal_income)
-
-original_lmmodel <- lm(time_spent_vigorous_exercise_7d ~ ., #+
-              # age:highest_educational_attainment +
-              # age:sex +
-              # sex:personal_income,
-                          # age+ 
-                          # highest_educational_attainment+
-                          # personal_income+
-                          # smoked_hundred_cigarettes+
-                          # # perceived_work_stress+
-                          # # illicit_drug_use+
-                          # health_region +
-                          # num_alc_drank_12m,
-              data = analysis_data_w_dummies)
-
-original_lmmodel_not_transformed <- lm(time_spent_vigorous_exercise_7d ~ ., #+
[…]

### 2024-12-03 scripts/08-health_region_map.R

@@ -1,65 +1,0 @@
-
-library(sf)
-library(tmap)
-library(tidyverse)
-health_region_shp <- st_read(
-  "../data/03-map_data/Health Regions Ontario Shapes/HR_035a18a_e.shp",
-  quiet = T)
-
-# tm_shape(canada_shp) +
-#   tm_borders() +
-tmap_mode('plot')
-
-HR_UID_top <- tibble(HU_ID = top_5_regions) |>
-  mutate(health_region = case_when(
-    HU_ID == 35951 ~ 'City of Ottawa Health Unit',
-    HU_ID == 35953 ~ 'Peel Regional Health Unit',
-    HU_ID == 35960 ~ 'Simcoe Muskoka District Health Unit',
-    HU_ID == 35970 ~ 'York Regional Health Unit',
-    HU_ID == 35995 ~ 'City of Toronto Health Unit'
-  ))
-
-HR_UID_btm <- tibble(HU_ID = bottom_5_regions) |>
-  mutate(health_region = case_when(
-    HU_ID == 35931 ~ 'Elgin-St Thomas Health Unit',
-    HU_ID == 35935 ~ 'Haliburton, Kawartha, Pine Ridge District ',
-    HU_ID == 35940 ~ 'Chatham-Kent Health Unit',
-    HU_ID == 35942 ~ 'Lambton Health Unit',
-    HU_ID == 35956 ~ 'Porcupine Health Unit'
-  ))
-
-
-
-HR_UID_top <- paste0(substr(top_5_regions, 1, 2), substr(top_5_regions, 4, 5))
-
-HR_UID_btm <- paste0(substr(bottom_5_regions, 1, 2), substr(bottom_5_regions, 4, 5))
-
-top_health_regions_shp <- health_region_shp |>
-  filter(HR_UID %in% HR_UID_top)
-
-btm_health_regions_shp <- health_region_shp |>
-  filter(HR_UID %in% HR_UID_btm)
-
-
-health_region_map <- tm_basemap('CartoDB.Positron') +
-  tm_shape(health_region_shp) +
-  tm_borders(colour = 'ENGNAME',
-             legend.show = F,
-             popup.vars = c('Name: '='ENGNAME'),
-             id = 'ENGNAME') +
-  tm_shape(top_health_regions_shp) +
-  tm_polygons(fill='indianred4',
-              fill_alpha = 0.5) +
-  tm_shape(btm_health_regions_shp) +
-  tm_polygons(fill='skyblue1',
[…]

### 2024-12-04 paper/paper.qmd
Change of map chunk for paper.qmd
Changed script from 08-health_region.R to 05-health_region.R

@@
 ---
-title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers respectively"
+title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers"
 subtitle: "Women are engaging in lesser physical activities than men.
@@
   - Chris Yong Hong Sen
-thanks: "Code and data are available at: [https://github.com/Monoji77/physical_health_ontario](https://github.com/Monoji77/physical_health_ontario)."
+thanks: "Code and data are available at: [https://github.com/Monoji77/Alcohol_Use_Ontario](https://github.com/Monoji77/Alcohol_Use_Ontario)."
 date: today
@@
 Additionally, age, highest educational attainment, and health region also play significant roles in determining physical activity levels.
-These findings highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing helpful information for designing targeted public health interventions to promote physical activity."
+These insights highlight the complex interplay of gender, lifestyle behaviors, and sociodemographic factors, providing valuable information for designing targeted public health interventions to promote physical activity."
 format: 
@@
 This is especially concerning as physical inactivity is a leading contributor to chronic diseases such as heart disease, diabetes, and mental health disorders.
-Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play an important role in shaping physical activity behaviors.
-Understanding these factors and their complex interactions is important for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.
+Factors such as age, sex, educational attainment, alcohol consumption, smoking habits, and regional disparities in health access play a critical role in shaping physical activity behaviors.
+Understanding these factors and their complex interactions is crucial for developing targeted interventions that can improve health outcomes and reduce health inequalities in Ontario.
 
@@
 Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper.
-At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered, and in [Appendix -@sec-survey_methodology], it contains a section regarding the survey and it's methodology.
+At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered.
 
@@
 Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow].
-Shape date was read into R using `sf` package, and generation of maps were done using functions in `tmap` package [@citeSF; @citationTmap].
+Generation of maps were done using functions in `tmap` package [@citationTmap].
 Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].
@@
 The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada.
-The goal of the CCHS survey is to collect reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
+The goal of the CCHS survey is to collect comprehensive and reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
 It aims to provide timely, accessible, and flexible data for monitoring health trends, studying small populations and rare characteristics, and addressing emerging health issues, ultimately helping improve the health and well-being of Canadians.
@@
 According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents).
-Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected in @fig-barplot-sex.
+Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected ing @fig-barplot-sex.
 ```{r sex predictor barplot}
@@
 
-Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain understanding in the relationship between our predictors and the response as discussed in @sec-discussion.
+Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into relationship between our predictors and the response as discussed in @sec-discussion.
 

### 2024-12-08 paper/paper.qmd
Final updates to paper
1. Shortened subtitle.
2. Corrected grammar and spelling.
3. Added more weaknesses of paper.
4. Added to telegraphing paragraph to include appendix B.
5. Added justification to 'Variable 1: Age' section as to why age would be included in model exploration.
6. Changed 'Histogram of', 'Boxplot of' 'Barplot of' to other terminology.

@@
 ---
-title: "Men and past smokers from ontario are potentially more physically active than women and non-smokers"
-subtitle: "Women are engaging in lesser physical activities than men.
-Furthermore, individuals who smoked before are more likely to engage in physical activity.
-Consumption of stimulant-related drugs could have also increased as well, resulting in increased overall physical activities in Ontario."
+title: "Men and past smokers from Ontario are potentially more physically active than women and non-smokers"
+subtitle: "Physical activity is discovered to be strongly associated with gender, smoking habits, and drug consumption."
 author: 
@@
 date-format: long
-abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographic and substance abuse.
+abstract: "This study explores factors influencing physical activity levels in Ontario, focusing on demographics and substance abuse.
 Our findings reveal that women engage in less physical activity than men, while individuals who have smoked in the past and those consuming stimulant-related drugs are more likely to be physically active.
@@
 
-Recent studies suggest that certain demographic groups, such as older adults and individuals with lower educational levels, are more likely to experience barriers to physical activity, while alcohol consumption and smoking are associated with lower physical activity levels [@citationBaumanPhysicalActivity].
+Recent studies suggest that certain demographic groups, such as older adults and individuals with lower educational levels, are more likely to experience barriers to physical activity.
+At the same time, alcohol consumption and smoking are associated with lower physical activity levels [@citationBaumanPhysicalActivity].
 However, limited research has explored how these factors specifically interact in Ontario, a province with diverse populations and health regions.
-As the burden of chronic disease grows, addressing these disparities becomes even more urgent [@citationWarburtonPhysicalActivity] .
+As the burden of chronic disease grows, addressing these disparities becomes even more urgent [@citationWarburtonPhysicalActivity].
 
@@
 @sec-data would provide a sense of the chosen Computing in the Humanities and Social Sciences dataset as well as data exploration of interested variables.
-@sec-Model goes into the details of the linear regression model fitted on the predictors of interest variables, for which it's results would be showcased in @sec-results.
-Finally @sec-discussion would interpret the results of the fitted linear regression model, which includes limitations of the paper.
-At the end of the paper in [Appendix -@sec-diagnostic_checks], this contains diagnostic checks for all models considered.
+@sec-Model goes into the details of the linear regression model fitted on the predictors of interest variables, for which its results would be showcased in @sec-results.
+@sec-discussion would then interpret the results of the fitted linear regression model, which includes the limitations of the paper.
+In [Appendix -@sec-diagnostic_checks], it contains diagnostic checks for all models considered.
+Finally, [Appendix -@sec-survey_methodology] goes through the background and surveying methodology of the CCHS dataset.
 
@@
 Reading of raw data and data manipulation was performed using `Tidyverse` package [@citeTidyverse] and parquet file handling was done using functions in `arrow` package [@citeArrow].
-Generation of maps were done using functions in `tmap` package [@citationTmap].
+Generation of maps was done using functions in `tmap` package [@citationTmap].
 Any other visual plots were using `ggplot`-related functions in Tidyverse [@citeTidyverse] and tables were generated using `kable` function in `knitr` package [@citeKnitR].
@@
 
-The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the university of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada.
+The dataset chosen is obtained from Computing in the Humanities and Social Sciences/CHASS at the University of Toronto [@SDA_CHASS], specifically the Canadian Community Health Survey (CCHS) annual component 2017-2018 which was collected by @StatisticsOfCanada.
 The goal of the CCHS survey is to collect comprehensive and reliable health-related data on the Canadian population to support health surveillance, research, and program evaluation.
@@
 
-Other datasets explored include sources from World Health Organisation and ParticipACTION.
+Other datasets explored include sources from the World Health Organisation and ParticipACTION.
 World Health Organisation provides aggregated data informing certain indicators of interest such as obesity levels and alcohol consumption.
 This can be found in the data section for Canada[@OtherSource1].
-Also, ParticipACTION provides key statistics regarding proportion of adult population in canada meeting the national guidelines of 150 minutes of moderate to vigorous physical activites, as well as how sedentary canadians are [@OtherSource2].
-However, the availability of raw individual level data is absent from these sources.
-Although the aggregated data gives us a sensing about general Canadian health, we are not able to focus our attention to the Ontario population.
-The dataset chosen from Canadian Community Health Survey (CCHS) annual component is ideal as it contains individual level data, where we are able to obtain the province the respondents are from.
+Also, ParticipACTION provides key statistics regarding the proportion of the adult population in Canada meeting the national guidelines of 150 minutes of moderate to vigorous physical activities, as well as how sedentary Canadians are [@OtherSource2].
+However, the availability of raw individual-level data is absent from these sources.
+Although the aggregated data gives us a sense of general Canadian health, we are not able to focus our attention on the Ontario population.
+The dataset chosen from the Canadian Community Health Survey (CCHS) annual component is ideal as it contains individual-level data, where we are able to obtain the province the respondents are from in order to obtain our desired target population.
 
@@
 
-The dataset provides 113,290 individual response from Canadians all throughout the country.
-Surveys were conducted to collect health-related data from respondents.
-The relevant data collected includes...
+The dataset provides 113,290 individual responses from Canadians throughout the country. Surveys were conducted to collect health-related data from respondents. The relevant data collected includes...
 
@@
 
-This dataset is highly suitable in determining both quantitative factors such as  c qualitative factors
+This dataset is highly suitable for determining both quantitative factors such as  c qualitative factors.
 
@@
 According to the questionnaire, respondents were asked in a phone survey to indicate whether they had engaged in any strenuous activity that lasted at least 10 minutes and made them sweat and breathe a little harder in the past 7 days.
-Afterwards, they were asked to indicate the total time spent on those activities.
-To address the concern of recall bias where an individual may not recall the specific timing in engaging with physical activity, surveyors sectioned physical activity into commuting, sports, work and volunteering.
+Afterward, they were asked to indicate the total time spent on those activities.
+To address the concern of recall bias where an individual may not recall the specific timing in engaging with physical activity, surveyors sectioned physical activity into commuting, sports, work, and volunteering.
 Then within each section, surveyors asked which day (Monday to Sunday) in the past week had they engaged in such activities that made them sweat and breathe a little harder.
@@
 
-For socio-demographic factors possibly affecting time spent on physical activites, the surveyors asked the respondent for their age (DHH_AGE) from 0 years old to 121 years old), sex (DHH_SEX) being either male or female, personal income obtained from tax forms submitted to the Canada Revenue Agency upon consent from respondent, geographical health region (ADM_Q037) that was provided by respondents when they agreed to the phone survey, and highest educational attainment (EHG2_04) ranging from less than high school diploma to master's degree and above.
-
-For substance abuse, respondents were asked how frequent they consumed alcoholic drinks in the past 12 months (ALC_Q015), ranging from less than once per month to everyday.
-Surveyors also asked respondents for usage of drugs including one-time marijuana or hashish in the past 12 months (DRG_Q005 to DRG_Q075), and whether they have smoked hundred cigarettes in the past 12 months (SMK_Q020).
+For socio-demographic factors possibly affecting time spent on physical activities, the surveyors asked the respondent for their age (DHH_AGE) from 0 years old to 121 years old), sex (DHH_SEX) being either male or female, personal income obtained from tax forms submitted to the Canada Revenue Agency upon consent from the respondent, geographical health region (ADM_Q037) that was provided by respondents when they agreed to the phone survey, and highest educational attainment (EHG2_04) ranging from less than high school diploma to master's degree and above.
+
+For substance abuse, respondents were asked how frequently they consumed alcoholic drinks in the past 12 months (ALC_Q015), ranging from less than once per month to every day.
+Surveyors also asked respondents about the usage of drugs including one-time marijuana or hashish in the past 12 months (DRG_Q005 to DRG_Q075), and whether they have smoked hundred cigarettes in the past 12 months (SMK_Q020).
 The last 2 questions allowed respondents to indicate a simple yes or no flag to express their answer.
 
-For health factors, questions were asked to determine an individual's body mass index ranging from underweight to obese and perceived life/work stress ranging from not at all stressful to extremely stressful.
+For health factors, questions were asked to determine an individual's body mass index ranging from underweight to obese, and perceived life/work stress ranging from not at all stressful to extremely stressful.
 Additionally, surveyors asked respondents if they have mood disorders (E.g.
-depression, bipolarism, mania), for which respondents replied with a simple yes or no.
-
-Overall, each entry obtained by respondents assigns quantitative values to each factor and the the response for which we can explore subsequently using linear regression to determine which factors have greater influence in the time spent on physical activities amongst active individuals.
+depression, bipolar, mania), to which respondents replied with a simple yes or no.
+
+Overall, each entry obtained by respondents assigns quantitative values to each factor and the the response for which we can explore subsequently using linear regression to determine which factors have greater influence on the time spent on physical activities amongst active individuals.
 
@@
 
-Since we are interested in individuals that are already engaging in strenuous activities, we can filter $time\_spent\_vigorous\_exercise\_7d$ to exclude 0.
+Since we are interested in individuals who are already engaging in strenuous activities, we can filter $time\_spent\_vigorous\_exercise\_7d$ to exclude 0.
 After removing such entries, we obtain the following summary statistics as shown in @tbl-summarised_statistics_time_spent_vigorous_exercise_7d.
@@
 
-Interestingly, the maximum time spent on physical activities the past 7 days was 9902 minutes, and this averages to 1415 minutes per day, which is equivalent to 23.5 hours in a day.
+Interestingly, the maximum time spent on physical activities in the past 7 days was 9902 minutes, and this averages to 1415 minutes per day, which is equivalent to 23.5 hours in a day.
 This is very unlikely as this suggests that an individual is constantly engaging in physical strenuous activity without sleeping.
@@
 #| label: fig-boxplot_time_spent_vigorous_exercise_7d
-#| fig-cap: Boxplot of time spent on physical activities in the past 7 days
+#| fig-cap: Summary statistic of time spent on physical activities in the past 7 days
 
@@
 
-By observing the histogram generated in @fig-hist-time_spent_vigorous_exercise_7d, we obtain a severely right skewed response histogram where most observations lie between 0 to 500 minutes.
+By observing @fig-hist-time_spent_vigorous_exercise_7d, we obtain a severely right-skewed response where most observations lie between 0 to 500 minutes.
 For the building of linear models in subsequent sections of the paper, this would not be ideal as linear regression assumes a normal distribution of the response variable, suggesting that a transformation would be required to fit a linear regression model.
@@
 #| label: fig-hist-time_spent_vigorous_exercise_7d
-#| fig-cap: Histogram of time spent on physical activities the past 7 days
+#| fig-cap: Time spent on physical activities the past 7 days
 
@@
      main='',
-     xlab = '',
+     xlab = 'no. of minutes',
      ylab = 'no. of individuals',
@@
 
-There are a total of 13 predictors of interest that is seemingly related to the time spent on physical activities by an individual in Ontario.
+There are a total of 13 predictors of interest that are seemingly related to the time spent on physical activities by an individual in Ontario.
 
 ### Variable 1: Age
-This is a variable of interest because age is seemingly correlated to The age groups of respondents were obtained and they exist in bins that spans ages 18 years to 74 years old.
+This is a variable of interest because age is seemingly correlated to time spent on physical activities.
+It is well known that the younger schooling population has more time and physical capacity to exercise than a working adult.
+In the International Journal of Behavioral Nutrition and Physical Activity, @citationUofExerciseAndAge discovered that younger individuals were exercising more than their older counterparts.
+
+The age groups of respondents were obtained and they exist in bins that span ages 18 years to 74 years old.
 
@@
 #| label: fig-barplot-age
-#| fig-cap: Histogram of time spent on physical activites the past 7 days
+#| fig-cap: Time spent on physical activities the past 7 days
 
@@
 
-From @tbl-summarised_statistics_age and @fig-boxplot-age, we are able to see that the the median age bin belongs to age group 40-44 years old.
+From @tbl-summarised_statistics_age and @fig-boxplot-age, we are able to see that the median age bin belongs to the age group 40-44 years old.
 We are able to see that there are no outlier values from @fig-boxplot-age, which means that the survey has obtained an acceptable number of respondents across all age bins.
@@
 #| label: fig-boxplot-age
-#| fig-cap: Boxplot of respondent ages
+#| fig-cap: Summary statistics of respondent ages
 
@@
 Sex is a possible influential predictor for total time spent on physical activity.
-Historically, men are associated with exercise and involved in more labour intensive roles such as construction which involves heavy lifting.
-According to @womenExerciseWHO, women are less active than men by 5% since 2000.
-However, with the advent of female only establishments such as female gyms, and 32% increase in memberships at fitness and health clubs from 2010 to 2019, it would be interesting to see if the sex of an individual still affects the time spent on physical activites among physically active individuals.
+Historically, men are associated with exercise and involved in more labour-intensive roles such as construction which involves heavy lifting.
+According to @womenExerciseWHO, women have been less active than men by 5% since 2000.
+However, with the advent of female-only establishments such as female gyms, and a 32% increase in memberships at fitness and health clubs from 2010 to 2019, it would be interesting to see if the sex of an individual still affects the time spent on physical activities among physically active individuals.
 
 According to @fig-barplot-sex, there is roughly an equal proportion of sexes from our dataset, with slightly more males (6082 respondents) than females (5859 respondents).
-Since this was a health survey conducted over phone call, there is an equal chance of a respondent to be male or female, which is roughly reflected ing @fig-barplot-sex.
+Since this was a health survey conducted over a phone call, there is an equal chance of a respondent being male or female, which is roughly reflected in @fig-barplot-sex.
 ```{r sex predictor barplot}
@@
 #| label: fig-barplot-sex
-#| fig-cap: Barplot of sex
+#| fig-cap: Frequency of sex variable
 data_cleaned_1 |>
@@
 
-In 2023, @EducationAndActiveLifestyle has done a study to show that highest educational attainment may be a leading factor for physical activity.
-Therefore, it would be worth exploring the effects of highest levels of education with the time spent on physical activities among the Ontario residents.
+In 2023, @EducationAndActiveLifestyle did a study to show that the highest educational attainment may be a leading factor for physical activity.
+Therefore, it would be worth exploring the effects of the highest levels of education on the time spent on physical activities among Ontario residents.
 
 According to @tbl-proportion_highest_educational_attainment, respondents who did not graduate secondary school account for only 5% of the population, and respondents who have at least a bachelor's degree account for 72% of the population.
-According to @fig-barplot_highest_educational_attainment, there seems to be class imbalance for this variable.
-
-However, as seen from post-secondary institution enrollment data from @statisticaPostSecEnrollment, there has been an increase in total number of Canadians enrolled in post secondary education from 2000 to 2020.
-Accounting for population changes from 2000 [@statCanadaPopulation2000] to 2022 [@statCanadaPopulation2022], there is still a real increase in the number of real enrollments into post secondary education.
-Therefore, it is very possible that this is representative of the Ontario's population where majority of the population has at least post secondary education.
+According to @fig-barplot_highest_educational_attainment, there seems to be a class imbalance for this variable.
+
+However, as seen from post-secondary institution enrollment data from @statisticaPostSecEnrollment, there has been an increase in the total number of Canadians enrolled in post-secondary education from 2000 to 2020.
+Accounting for population changes from 2000 [@statCanadaPopulation2000] to 2022 [@statCanadaPopulation2022], there is still a real increase in the number of real enrollments in post-secondary education.
+Therefore, it is very possible that this is representative of Ontario's population where the majority of the population has at least post-secondary education.
 
@@
 #| label: fig-barplot_highest_educational_attainment
-#| fig-cap: Barplot of highest_educational_attainment
+#| fig-cap: Frequency of highest_educational_attainment variable
 data_cleaned_1 |>
@@
 
-Health region is a categorical variable which tells us the geographical location an individual belongs to defined by Provincial Ministries of Health (as detailed by documentation provided by CHASS).
+Health region is a categorical variable that tells us the geographical location an individual belongs to defined by Provincial Ministries of Health (as detailed by documentation provided by CHASS).
 It could be possible that certain health regions have more funding that are projected towards increasing awareness towards having an active lifestyle.
@@
 @fig-map_health_region shows us the various health regions in Ontario.
-The blue regions shows us health regions for which lesser respondents belong to while red regions show us health regions for which more respondents belong to.
+The blue regions show us health regions to which fewer respondents belong while the red regions show us health regions to which more respondents belong.
 
@@
 
-By using a GIS focused library in R called tmap [@citationTmap], we are able to generate the map shown in @fig-map_health_region.
+By using a GIS-focused library in R called tmap [@citationTmap], we are able to generate the map shown in @fig-map_health_region.
 
@@
 Overall, @fig-barplot_health_region shows us the actual number of observations belonging to the top five and bottom 5 occurring health regions.
-We can see that most respondents come from the City of Ontario Health Unit (Health_Region = 35995) and lesser respondents belong to Porcupine Health Unit (Health_Region = 35956).
+We can see that most respondents come from the City of Ontario Health Unit (Health_Region = 35995) and fewer respondents belong to the Porcupine Health Unit (Health_Region = 35956).
 
@@
 #| label: fig-barplot_health_region
-#| fig-cap: Barplot of top and bottom five occuring health regions
+#| fig-cap: Frequency of top and bottom five occuring health regions
 
@@
 
-In the dataset, personal income (INCDGPER) is a categorical variable that places an individuals annual income into a range of income values.
+In the dataset, personal income (INCDGPER) is a categorical variable that places an individual's annual income into a range of income values.
 Below summarises the encoded number and the corresponding meaning of each encoded number.
@@
 #| label: fig-barplot_personal_income
-#| fig-cap: Barplot of personal income
+#| fig-cap: Frequency of personal income
 
@@
 
-From @fig-barplot_frequency_alcohol_consumption, we are able to see that most respondents from this dataset drink about 2-3 times per week (2973 respondents) as seen from the red bar.
+From @fig-barplot_frequency_alcohol_consumption, we can see that most respondents from this dataset drink about 2-3 times per week (2973 respondents) as seen from the red bar.
 Also, we are seeing fewer respondents drinking about 4-6 times per week (887 respondents) as seen from the blue bar.
@@
 #| label: fig-barplot_frequency_alcohol_consumption
-#| fig-cap: Barplot of frequency of alcohol consumption
+#| fig-cap: Frequency of alcohol consumption
 max_alc <- data_cleaned_1 |> 
@@
 
-This is a binary variable which indicates whether an individual has used drugs in the past 12 months prior to the interview.
+This is a binary variable that indicates whether an individual has used drugs in the past 12 months prior to the interview.
 Drugs include marijuana, hashish, cocaine, amphetamines, speed, methamphetamine, crystal meth, ecstasy, hallucinogens, and sniffing glue, gasoline or other solvents.
@@
 
-It would be interesting to see if drug consumption has an effect on physical activeness.
+It would be interesting to see if drug consumption affects physical activity.
 @PhysicalActivityAndSubstanceAbuse exerts that individuals who are more active tend to have lower substance abuse, suggesting a correlation between these two.
-Having this variable would be interesting to explore the relationship of an individuals drug usage and their time spent on physical activities.
-
-@fig-barplot_illicit_drug_use shows us the number of respondents who have (or haven't) consumed drugs 12 months prior to their interviews.
-We can see that majority of respondents have not consumed drugs in the given timeframe.
-However, from @tbl-illicit_drug_use, we can see that there are still a number of respondents who had consumed drugs in the past 12 months, particularly a fifth of the sampling population.
-According to @SubstanceUseAndAddiction, they estimate that about 21% of canadians, or 6 million people, will meet the criteria for substance use in their lifetime, and this is inclusive of alcohol and that there is a growing opiod crisis in Ontario.
-Therefore the obtained sample is somewhat representative of usage of drugs among the Ontario population.
+Having this variable would be interesting to explore the relationship between an individual's drug usage and their time spent on physical activities.
+
+@fig-barplot_illicit_drug_use shows us the number of respondents who have (or haven't) consumed drugs 12 months before their interviews.
+We can see that the majority of respondents have not consumed drugs in the given timeframe.
+However, from @tbl-illicit_drug_use, we can see that there are still several respondents who had consumed drugs in the past 12 months, particularly a fifth of the sampling population.
+According to @SubstanceUseAndAddiction, they estimate that about 21% of Canadians, or 6 million people, will meet the criteria for substance use in their lifetime, and this is inclusive of alcohol and there is a growing opioid crisis in Ontario.
+Therefore the obtained sample is somewhat representative of the usage of drugs among the Ontario population.
 ```{r Illicit drug use summary table}
@@
 #| label: fig-barplot_illicit_drug_use
-#| fig-cap: Barplot of illicit drug use the past 12 months
+#| fig-cap: Illicit drug use the past 12 months
 data_cleaned_1 |>
@@
 Respondents were asked whether they had smoked more than 100 cigarettes in their lifetimes.
-This variable is therefore binary, with '1' representing respondents that meet the critera and '0' otherwise.
+This variable is therefore binary, with '1' representing respondents that meet the criteria and '0' otherwise.
 
 It is well known that excessive smoking has negative health impacts on individuals.
-A cross sectional study conducted by @SmokingAndPhysicalActiveness found that smokers are less likely to be physically active compared to non-smokers.
-Hence, it would be interesting to discover if there is an effect of smoking more than 100 cigarettes on the time spent on physical activites.
-
-@fig-barplot-cigarettes shows us that within the sample, the visual difference between the number of respondents who smoked more than 100 cigarettes (blue column) and respondents who smoked lesser than 100 cigarettes are not that big.
+A cross-sectional study conducted by @SmokingAndPhysicalActiveness found that smokers are less likely to be physically active compared to non-smokers.
+Hence, it would be interesting to discover if there is an effect of smoking more than 100 cigarettes on the time spent on physical activities.
+
+@fig-barplot-cigarettes shows us that within the sample, the visual difference between the number of respondents who smoked more than 100 cigarettes (blue column) and respondents who smoked less than 100 cigarettes is not that big.
 Specifically, 57% of respondents have smoked less than 100 cigarettes while the remaining 43% of respondents smoked more than 100 cigarettes.
-As the classes in this variable is almost split half,
+As the classes in this variable are almost split in half.
 
@@
 
-@fig-pyrmaid-cigarettes shows us cigarette consumption by age bins among out sample population.
-As respondent age increases from the first bin (18 to 19 year old) to the fourth bin (30 to 34 year old), we can see the number of respondents who have smoked more than 100 cigarettes in their lifetime increased from less than 125 respondents to more than 500 respondents (about 3 times more respondents between the two age groups).
+@fig-pyrmaid-cigarettes shows us cigarette consumption by age bins among our sample population.
+As respondent age increases from the first bin (18 to 19 years old) to the fourth bin (30 to 34 years old), we can see the number of respondents who have smoked more than 100 cigarettes in their lifetime increased from less than 125 respondents to more than 500 respondents (about 3 times more respondents between the two age groups).
 ```{r cigarette consumption pyramid figure}
@@
 #| label: fig-pyrmaid-cigarettes
-#| fig-cap: Pyramid of cigarettes consumption in respondent's lifetime by age groups
+#| fig-cap: Cigarette consumption in respondent's lifetime by age groups
 
@@
 
-Let $k$ be the individual from CCHS dataset.
+Let $k$ be the individual from the CCHS dataset.
 Define $y_k$ as the time spent on physical activities for individual $k$.
-Then $Age_k$, $Sex_k$, $Educ_k$ and $Alc_k$ represents individual i's age, sex, highest educational attainment, and frequency of alcohol consumption the past 12 months respectively.
+Then $Age_k$, $Sex_k$, $Educ_k$, and $Alc_k$ represent individual i's age, sex, highest educational attainment, and frequency of alcohol consumption in the past 12 months respectively.
 Also, $Drug_k$ and $Cigar_k$ would be indicator explanatory variables derived from the binary variables in our dataset representing if individual k had consumed drugs in the past 12 months and if he/she has smoked 100 cigars in their lifetime respectively.
 Finally, $HealthRegion_k=35953$ and $HealthRegion_k=35970$ are also indicator explanatory variables derived from the categorical variable health_region in our dataset, which represents if individual k belongs to Health Region 35953 (Peel Regional Health Unit) and if he/she belongs to Health Region 35970 (York Regional Health Unit).
-Note that the reference with respect to Health Region would be individuals that are not in both health region 35953 and 35970.
+Note that the reference concerning Health Region would be individuals that are not in both Health Region 35953 and 35970.
 
@@
 Since health region is a categorical variable with no implied order, To handle health region, we will be creating dummy variables for each health region, specifically $I*(HealthRegion_k=i)$ for $i \in {health\_region}$, and the reference chosen for health region would be the health_region = 35926 (The District of Algoma HU).
-We define $Income_k$ to be the the personal income of individual k in our dataset.
+We define $Income_k$ to be the personal income of individual k in our dataset.
 
@@
 
-We then conducted model reduction using backward feature elimination using Bayesian Information Criterion (BIC) as our scoring metric to obtain the least number of predictors that minimises the residual errors from the reduced model.
+We then conducted model reduction using backward feature elimination using the Bayesian Information Criterion (BIC) as our scoring metric to obtain the least number of predictors that minimizes the residual errors from the reduced model.
 This model reduction was performed using `stats` package from the @citeR as well.
@@
 @tbl-BIC_values shows us the BIC values for both the full model and the reduced model.
-After removing certain explanatory variables, the BIC score has decreased, indicating that the reduced model is has stronger explanatory power while having a less complex model.
+After removing certain explanatory variables, the BIC score has decreased, indicating that the reduced model has stronger explanatory power while having a less complex model.
 Therefore, the reduced model is chosen from the original full model.
@@
 
-Overall, the individual t-test for our predictors were found to be below our set threshold of 95% alpha.
+Overall, the individual t-test for our predictors was found to be below our set threshold of 95% alpha.
 The confidence interval for each predictor is listed in @tbl-confidence_interval_lmmodel shown below.
@@
 
-## Males are more likely to engage in physical activities{#sec-men_physical_activites}
-
-According to @fig-physical_activity_by_sex, we can observe that amongst individuals that are active, men spend (on average) more time on physical activity than women.
-In particular, according to our sample men spend on average 639 minutes a week on physical activites while women spend on average 437 minutes a week on physical activities.
+## Males are more likely to engage in physical activities{#sec-men_physical_activities}
+
+According to @fig-physical_activity_by_sex, we can observe that amongst active individuals, men spend (on average) more time on physical activity than women.
+In particular, according to our sample men spend on average 639 minutes a week on physical activities while women spend on average 437 minutes a week on physical activities.
 
@@
 #| label: fig-physical_activity_by_sex
-#| fig-cap: "Bar plot of average time spent on physical activity by sex"
+#| fig-cap: Average time spent on physical activity by sex
 #| warning: false
@@
 This means that if an individual is a male, this increases the time spent on physical activities by 26%.
-While we were expecting increase in the overall engagement in physical activity among women as explained in @sec-sex_exploration, men are still more likely to engage in physical activities than women.
+While we were expecting an increase in the overall engagement in physical activity among women as explained in @sec-sex_exploration, men are still more likely to engage in physical activities than women.
 In 2018, Global News Canada reported that 31% of women and 25% of men have insufficient physical activity [@citationMenVsWomenExercise].
@@
 
-In fact, in a study published in the Journal of American College of Cardiology, they discovered that women who were physically active would experience greater health benefits than men who were also physically active [@citationWomenSurviveBetterExercise].
+In fact, in a study published in the Journal of American College of Cardiology, they discovered that physically active women would experience greater health benefits than men who were also physically active [@citationWomenSurviveBetterExercise].
 Therefore, this means that more can be done to address the gender gap in the engagement of physical activities.
@@
 From elementary school, Ontario students are exposed to health education where schools would educate about the harms of smoking.
-It is well known that smoking has negative effect's on an individual's health.
+It is well known that smoking has negative effects on an individual's health.
 Therefore it is unexpected that the linear regression model showed a positive coefficient for the indicator variable for 'having smoked 100 cigarettes' in an individual's lifetime.
 
-The linear model is telling us that if an individual has smoked 100 cigarettes in their lifetime, they are 13% more likely to spend more time being physically active than an individual who has not smoked more than 100 cigarettes in the same time frame.
+The linear model tells us that if an individual has smoked 100 cigarettes in their lifetime, they are 13% more likely to spend more time being physically active than an individual who has not smoked more than 100 cigarettes in the same time frame.
 While it may seem contradicting to what we believe, there may be merit to this value.
 From @fig-physical_activity_by_smoking_status, we can see that individuals who had smoked 100 cigarettes in their lifetime spent more time on physical activity than individuals who have not smoked 100 cigarettes.
-In particular, our data shows that on average, individuals who smoked 100 cigarettes spend 618 minutes on physical activities in a week while those who don't spend only 418 minute on physical activities.
+In particular, our data shows that on average, individuals who smoked 100 cigarettes spend 618 minutes on physical activities in a week while those who didn't spend only 418 minutes on physical activities.
 
@@
 #| label: fig-physical_activity_by_smoking_status
-#| fig-cap: "Bar plot of average time spent on physical activity by smoking status"
+#| fig-cap: Average time spent on physical activity by smoking status
 #| warning: false
@@
 
-A team from the school of Public Health Sciences at University of Waterloo had explored patterns and trends for tobbaco use in canada.
+A team from the School of Public Health Sciences at the University of Waterloo explored patterns and trends for tobacco use in Canada.
 What they found was that 70% of Canadians who had smoked before have quit as of 2020 @citationSmokingPatternsWaterlooUniversity.
-While smoking may have been affected an individual's physical capabilities, it could have also increased awareness regarding the importance of their health.
-Since most Canadians who had a history of smoking had quit smoking, it is very possible that they have greater health conscious.
+While smoking may have affected an individual's physical capabilities, it could have also increased awareness regarding the importance of their health.
+Since most Canadians who have a history of smoking have quit smoking, it is very possible that they have greater health consciousness.
 Coupled with increased government support for programs with the sole purpose of increasing physical activeness in Ontario [@citationGovernmentSupportExercise], it is no longer surprising why individuals who smoked 100 cigarettes in their lifetime are on average 13% more likely to engage in physical activities.
@@
 This means that if an individual from Ontario has consumed drugs listed in @sec-drug_exploration, they would spend on average 17% more time on physical activities.
-This is also another surprising finding since we have also discussed in @sec-drug_exploration that there is a negative correlation between drug use and physical activeness, but our findings for our Ontario population suggests a positive correlation.
+This is also another surprising finding since we have also discussed in @sec-drug_exploration that there is a negative correlation between drug use and physical activeness, but our findings for our Ontario population suggest a positive correlation.
 From @fig-physical_activity_by_illicit_drug_use, we can see that individuals who consumed drugs listed in @sec-drug_exploration in their lifetime spent more time on physical activity than individuals who have not consumed drugs.
-In particular, our data shows that on average, individuals are drug abusers spend 670 minutes on physical activities in a week while those who don't spend only 503 minute on physical activities.
+In particular, our data shows that on average, individuals who are drug abusers spend 670 minutes on physical activities in a week while those who don't spend only 503 minutes on physical activities.
 
@@
 #| label: fig-physical_activity_by_illicit_drug_use
-#| fig-cap: "Bar plot of average time spent on physical activity by illicit drug use"
+#| fig-cap: Average time spent on physical activity by illicit drug use"
 #| warning: false
@@
 Of the listed drugs, amphetamine and methamphetamine were identified as well.
-According to @citationDrugsAndActiveLifestyle, they mentioned that the use of meth (short term for methamphetamine) leads to increased concentrations of dopamine, which is a mood enhancer.
-This is coupled with the increase of body movement.
-Likewise, in a study conducted by the Indiana University [@citationAmphetamineandExercise], they found that the use of Amphetamine masks fatigue, allowing for individuals to exercise longer.
+According to @citationDrugsAndActiveLifestyle, they mentioned that the use of meth (short for methamphetamine) leads to increased concentrations of dopamine, which is a mood enhancer.
+This is coupled with the increase in body movement.
+Likewise, in a study conducted by Indiana University [@citationAmphetamineandExercise], they found that the use of Amphetamine masks fatigue, allowing individuals to exercise longer.
 This shows that it is possible that the inclusion of these drugs could have explained the surprising increase in time spent on physical activities despite the expectation that an individual who abuses drugs would have lower exercise.
 
-The inclusion of these drugs in the survey could explain the positive coefficient seen from our model.
+The inclusion of these drugs in the survey results in aggregation bias as many drugs which may have varying effects on physical activity were clumped into this single inclusion variable.
+This bias could explain the positive coefficient seen from our model.
 However, this is just a hypothesis and an experiment would need to be conducted to explore if this finding is true, where we obtain a sample of drug users who consumed amphetamine and methamphetamine and drug users who did not consume the two drugs.
@@
 Mitigating this requires complementary objective measures, such as wearable activity trackers or biomarkers.
+
+Furthermore, the intent of aggregating data to protect an individual's privacy in census data collection caused certain variables to be too over-generalized, and have limited categorizations for meaningful exploration.
+For instance, the highest educational attainment variable only has 3 bins, not having attained a secondary certification, having attained secondary certification, and having post-secondary certification.
+The first category includes both individuals that went to school and individuals that never went to school.
+Also, the last category includes individuals obtaining post-secondary diplomas, bachelor's, master's, and PhD degrees.
+Each category that was otherwise aggregated to fit the three categories during dataset collection might have had a different effect on the response variable, resulting in aggregation bias and potentially losing important information.
+
+Also, age has inconsistent bins, particularly for 18 to 19 year olds and every other bin, therefore this would result in fewer individuals that belong to the 18 to 19 year old bin.
+This would lead to biased medians when conducting summary statistics and undersampling due to a smaller inclusion criterion for this specific bin, resulting in decreased precision when conducting statistical tests like chi-square test.
+
 \newpage
@@
 
-Our response variable was log transformed to fulfill diagnostic checks for linear regression model.
-Observe the change in qq plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log transformed.
-From the left plot of @fig-diagnostic_plot_qq, we can see that values of standardised residuals are a lot larger than expected at higher quantile values.
-This suggests that the distribution of the residuals is severely right skewed.
+Our response variable was log-transformed to fulfill diagnostic checks for the linear regression model.
+Observe the change in the qq plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log-transformed.
+From the left plot of @fig-diagnostic_plot_qq, we can see that values of standardized residuals are a lot larger than expected at higher quantile values.
+This suggests that the distribution of the residuals is severely right-skewed.
 This contradicts the assumption of normality of residuals, which would make our linear regression model unreliable.
 
-From the right plot of @fig-diagnostic_plot_qq, after we transform our response using the natural logarithm function ($ln$), we can see that most of the residuals lie on the y=x line in the qqplot, suggesting that the sampling distribution is approximately normal.
+From the right plot of @fig-diagnostic_plot_qq, after we transform our response using the natural logarithm function ($ln$), we can see that most of the residuals lie on the y=x line in the qq plot, suggesting that the sampling distribution is approximately normal.
 
@@
 p2_1 <- plot(original_model,2, col=adjustcolor('black', alpha=0.2), sub='')
-title(main='log transformed')
-```
-
-Now, we observe the change in the scale location plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log transformed.
+title(main='log-transformed')
+```
+
+Now, we observe the change in the scale location plot of the fitted linear regression model before our response ($time\_spent\_on\_vigorous\_exercise\_7d$) was log-transformed and after it was log-transformed.
 From the left plot of @fig-diagnostic_plot_scale_location, we can see a weak but visible trend that as the fitted values increase, so does the square root of the standardized residuals.
-This suggests that the residuals are not randomly distributed, which goes against the assumption of linear regression that residuals are independent from each other.
+This suggests that the residuals are not randomly distributed, which goes against the assumption of linear regression that residuals are independent of each other.
 Furthermore, the mean red line is not horizontally flat.
-This is suggesting a positive trend.
-This tells us that the variance of the residuals is increasing, therefore implying that our residuals do not satisfy homoscedasticity assumption of linear regression, which would lead to an unreliable linear regression model.
-
-From the right plot of @fig-diagnostic_plot_scale_location, we can see that the standardised residuals are randomly scattered across the plot across all fitted values.
-This suggests that residuals are independant from each other.
-Furthermore, the mean red line is (almost) horizontally flat, for which suggests that the variance of residuals across all fitted values are constant.
-Therefore, homoscedasticity assumption for linear regression is also satisfied, allowing us to interpret the coefficient estimates of our fitted linear model with greater reliability.
+This suggests a positive trend.
+This tells us that the variance of the residuals is increasing, therefore implying that our residuals do not satisfy the homoscedasticity assumption of linear regression, which would lead to an unreliable linear regression model.
+
+From the right plot of @fig-diagnostic_plot_scale_location, we can see that the standardized residuals are randomly scattered across the plot across all fitted values.
+This suggests that residuals are independent from each other.
+Furthermore, the mean red line is (almost) horizontally flat, which suggests that the variance of residuals across all fitted values is constant.
+Therefore, the homoscedasticity assumption for linear regression is also satisfied, allowing us to interpret the coefficient estimates of our fitted linear model with greater reliability.
 
@@
 plot(original_model,3, sub='')
-title(main='log transformed')
+title(main='log-transformed')
 ```
@@
 By performing log transformation of my response, we have satisfied the linear regression assumptions, allowing us to interpret from a reliable linear regression model, therefore explaining the need to transform our response variable in @sec-Model.
-This would also allow us to perform model reduction outlined in @sec-model_justification
+This would also allow us to perform model reduction outlined in @sec-model_justification.
 
@@
 
-The final linear model was obtained after performing backward feature elimination using BIC score as the criterion.
-In this section, we will be showing that the final model satisfies linear regression assumptions.
-
-@fig-diagnostic_plot_qq_final_model shows us that the values of standardised residuals are approximately equal to the theoretical quantiles, which suggests that the residuals are normally distributed.
+The final linear model was obtained after performing backward feature elimination using the BIC score as the criterion. This section shows that the final model satisfies linear regression assumptions.
+
+@fig-diagnostic_plot_qq_final_model shows us that the values of standardized residuals are approximately equal to the theoretical quantiles, which suggests that the residuals are normally distributed.
 ```{r}
@@
 From @fig-diagnostic_plot_scale_location_final_model, we can see that the standardised residuals are randomly scattered across the scale location plot, which suggests that residuals are independant from each other.
-Furthermore the mean red line is (almost) horizontal which suggests that variance of residuals are constant across all fitted values.
-This means that the residuals obtained from the final model satisfies homoscedasticity assumption in linear regression
+Furthermore, the mean red line is (almost) horizontal which suggests that the variance of residuals is constant across all fitted values.
+This means that the residuals obtained from the final model satisfy the homoscedasticity assumption in linear regression.
 ```{r}
@@
 
-Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into relationship between our predictors and the response as discussed in @sec-discussion.
+Overall, since the residual assumptions in linear regression are satisfied, we are able to use the coefficient estimates obtained from our final fitted linear model to gain insights into the relationship between our predictors and the response as discussed in @sec-discussion.
 
@@
 
-The purpose of CCHS is to provide cross sectional estimates of factors related to different health determinants in Canada.
+The purpose of CCHS is to provide cross-sectional estimates of factors related to different health determinants in Canada.
 The use case for the Canadian health dataset is non-exhaustive and many have used it for the betterment of Canada.
-In the department of sociology and social work, University of Toronto, researchers have used the weight and weight perception data from the 2017-2018 CCHS to predict adolescents mental health, and they were able to find out that normal weighted girls are not immune from poor mental health arising from poor weight perceptions, hence urging specialists in the mental health field to increase support for these girls that have not been receiving enough avenues [@citationUofTCCHS].
-The Canadian Partnership Against Cancer also uses CCHS data to access the severity of the cervical cancer among Canadians, streamlining their cancer screening process as a result and urging the country to focus resources to address inequalities in access to cervical cancer prevention and treatment options between indigenous communities and poor regions [@citationCanadianPartnership2; @citationCanadianPartnership1].
+In the Department of Sociology and Social Work, University of Toronto, researchers have used the weight and weight perception data from the 2017-2018 CCHS to predict adolescents' mental health, and they were able to find out that normal weighted girls are not immune from poor mental health arising from poor weight perceptions, hence urging specialists in the mental health field to increase support for these girls that have not been receiving enough avenues [@citationUofTCCHS].
+The Canadian Partnership Against Cancer also uses CCHS data to assess the severity of cervical cancer among Canadians, streamlining their cancer screening process as a result and urging the country to focus resources on addressing inequalities in access to cervical cancer prevention and treatment options between Indigenous communities and poor regions [@citationCanadianPartnership2; @citationCanadianPartnership1].
 The one thing that ties many of the health-related studies arising from the use of CCHS studies is to address underrepresented communities in Canada.
 
-With the purpose of unifying data across health sources, the Statistics of Canada had began conducting surveys yearly since 2007.
-From 2015, there was a redesign to their survey methodology that is used for the dataset that was used in this paper.
+With the purpose of unifying data across health sources, the Statistics of Canada has begun conducting surveys yearly since 2007.
+Since 2015, there has been a redesign to the survey methodology that is used for the dataset that was used in this paper.
 
@@
 
-- individuals living on reserves, more commonly known as indigenous populations with their own governance within Canada ()
-- full time servicemen in the Canadian Armed Forces
+- individuals living on reserves, more commonly known as Indigenous populations with their own governance within Canada
+- full-time servicemen in the Canadian Armed Forces
 - youths aged 12 to 17 residing in foster homes
-- individuals residing in jail, nursing homes, mental health institutions and shelters
-- individuals in quebec health region of Nunavik and Terres-Cries-de-la-Baie-James
-
-According to CCHS, they account for lesser than 3% of the population
+- individuals residing in jail, nursing homes, mental health institutions, and shelters
+- individuals in the Quebec health region of Nunavik and Terres-Cries-de-la-Baie-James
+
+According to CCHS, they account for less than 3% of the population.
 
@@
 The sample for the CCHS is drawn from the respective sampling frames for adults and youth.
-Approximately 130,000 individuals are selected for the survey, with around 120,000 respondents allocated to the adult population and 10,000 to the youth population.
+Approximately 130,000 individuals were selected for the survey, with around 120,000 respondents allocated to the adult population and 10,000 to the youth population.
 The sample is stratified by health region, ensuring that each health region has a minimum of 500 respondents.
@@
 
-The Canadian Community Health Survey (CCHS) employs a systematic approach to address nonresponse at both the household and person levels.
+The Canadian Community Health Survey (CCHS) employs a systematic approach to address nonresponse at both the household and personal levels.
 This ensures that the survey results remain representative of the general population, even when certain individuals or households do not respond.
@@
 The selected individuals are then grouped into response homogeneity groups (RHGs) based on their calculated response probabilities.
-Within each RHG, an adjustment factor is calculated to correct for person-level.
+Within each RHG, an adjustment factor is calculated to correct for the personal level.
 
